### 2012-06-19 3dimpvt/_region_.tex
just waiting for some images

\message{ !name(paper.tex)}\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

% \cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{****} % *** Enter the 3DIMPVT Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
\ifcvprfinal\pagestyle{empty}\fi
\begin{document}

\message{ !name(paper.tex) !offset(204) }
\begin{figure}
  \centering
  \includegraphics[height=1.5in]{scoringFunction.pdf}
  \caption{For each tile, we want to select an image such that $d$ and $\alpha$ are minimized}
  \label{fig:scoringFunction}
\end{figure}

%\begin{figure}
%  \centering
%  \includegraphics[height=1.5in]{directMapping.png}
%  \caption{The result of direct texture mapping based on locally
%    optimized textures.}
%  \label{fig:directMapping}
%\end{figure}
\message{ !name(paper.tex) !offset(699) }

\end{document}