### 2012-06-29 3dimpvt/paper.tex

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{cvpr} \usepackage{times} \usepackage{epsfig}
\usepackage{graphicx} \usepackage{amsmath} \usepackage{amssymb}
\usepackage{subfig}

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first
% latex run, let it finish, and you should be clear).
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

% \cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{****} % *** Enter the 3DIMPVT Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in
% camera-ready
\ifcvprfinal\pagestyle{empty}\fi
\begin{document}

%%%%%%%%% TITLE
\title{Texture Mapping 3D Models of Indoor Environments with Noisy
  Camera Poses}

\author{First Author\\
  Institution1\\
  Institution1 address\\
  {\tt\small firstauthor@i1.org}
  % For a paper whose authors are all at the same institution, omit
  % the following lines up until the closing ``}''.  Additional
  % authors and addresses can be added with ``\and'', just like the
  % second author.  To save space, use either the email address or
  % home page, not both
  \and
  Second Author\\
  Institution2\\
  First line of institution2 address\\
  {\small\url{http://www.author.org/~second}} }

\maketitle
% \thispagestyle{empty}

%%%%%%%%% ABSTRACT
\begin{abstract}
  Automated 3D modeling of building interiors is useful in
  applications such as virtual reality and environment
  mapping. Applying textures to these models is an important step in
  generating photorealistic visualizations of data collected by
  modeling systems. Camera pose recovery in such systems often suffers
  from inaccuracies, resulting in visible discontinuties when images
  are projected adjacently onto a plane for texturing. We propose two
  approaches for reducing discontinuities in texture mapping 3D models
  made of planar surfaces. The first one is tile based and can be used
  for images and planes at arbitrary angles relative to each
  other. The second one, which we refer to as our seam minimization
  approach, results in a more seamless texture but is only applicable
  where camera axes for images are closely aligned with plane normals
  of the surfaces to be textured. The effectiveness of our approaches
  are demonstrated on two indoor datasets.
\end{abstract}

%%%%%%%%% BODY TEXT
\section{Introduction}
\label{sec:introduction}
Three-dimensional modeling of indoor environments has a variety of
applications such as training and simulation for disaster management,
virtual heritage conservation, and mapping of hazardous sites. Manual
construction of these digital models can be time consuming, and as
such, automated 3D site modeling has garnered much interest in recent
years.

The first step in automated 3D modeling is the physical scanning of
the environment's geometry. An indoor modeling system must be able to
recover camera poses within an environment while simulatenously
reconstructing the 3D structure of the environment itself
\cite{chen2010indoor, hz, kua2012loopclosure, liu2010indoor}.This is
known as the simultaneous localization and mapping (SLAM) problem, and
is generally solved by taking readings from laser range scanners,
cameras, and inertial measurement units (IMUs) at multiple locations
within the environment.

Mounting such devices on a platform carried by an ambulatory human
provides unique advantages over vehicular-based systems on wheels in
terms of agility and portability, but can also result in larger
localization error \cite{liu2010indoor}. As a result, common methods
for texture mapping generally produce poor results.

In this paper, we present a number of approaches to texture mapping 3D
models of indoor environments made of planar surfaces in the presence
of uncertainty and noise in camera poses. In particular, we consider a
human-operated backpack system with a number of laser range scanners
as well as 2 cameras facing left and right, each equipped with fisheye
lenses reaching an approximately 180$^{\circ}$ field of view and
taking photos at a rate of 5 Hz. With the cameras and laser scanners
active, the human operator wearing the backpack takes great care to
walk a path such that every wall in the desired indoor environment is
traversed and scanned lengthwise at least once, resulting in many
thousands of captured images.

Applying multiple localization and loop-closure algorithms on the data
collected by the ondoor sensors \cite{chen2010indoor,
  kua2012loopclosure, liu2010indoor}, the backpack is
localized\footnote{In this paper, we use the terms localization and
  pose recovery interchangeably, in that they both refer to recovering
  position and orientation.}  over its data collection period. This
requires recovering the 6 degrees of freedom for the backpack as well
as all its sensors, including the cameras rigidly mounted on it. Once
this is complete, the data from the laser range scanners is used to
generate a 3D point cloud of the surrounding environment and a 3D
planar model is generated \cite{sanchez2012point}. To do so, we
approximate normal vectors for each point in the point cloud by
gathering neighboring points within a small radius and processing them
through principal component analysis. These normal vectors allow for
the classification and grouping of adjacent points into structures
such as walls, ceilings, floors, and staircases. A RANSAC algorithm is
then employed to fit polygonal planes to these structured groupings of
points, resulting in a fully planar model. This model, consisting of
multiple 2D polygonal planes in 3D space, along with the set of images
captured by the backpack's cameras and their noisy 3D poses, can be
considered the input to our texture mapping problem.

This problem straddles the fields of image-based localization
refinement as well as image alignment and mosaicing, and thus our
proposed solution shares many similarities with past research in those
areas, which will be examined in Section \ref{sec:existingApproaches}. However, given our knowledge of approximate camera poses as well as the 3D geometry of the environment to be textured, we can more accurately and efficiently texture our planes.

The remainder of the paper is organized as follows. Section
\ref{sec:simpleTextureMapping} describes two variants of a tile-based
mapping approach and their shortcomings.  Section
\ref{sec:existingApproaches} discusses past research in related
fields, demonstrates two previous attempts at improving texture
alignment, and exhibits their inadequacies for our datasets. Section
\ref{sec:proposedApproach} presents our proposed seam minimization
approach to texture mapping. Section \ref{sec:resultsAndConclusions}
contains results and conclusions.

In all subsequent sections, we discuss the process of texture mapping
a single plane, as the texturing of each of our planes is independent
and can be completed in parallel.

\section{Tile-Based Texture Mapping}
\label{sec:simpleTextureMapping}
\begin{figure}
  \centering
  \includegraphics[height=2in]{Projection.pdf}
  \caption{Planes are specified in 3D space by four corners $C_1$ to
    $C_4$. Images are related to each plane through the camera
    matrices $P_{1..M}$. }
  \label{fig:projection}
\end{figure}

The geometry of the texture mapping process for a plane is shown in
Figure \ref{fig:projection}.  As described earlier, we are provided
with a set of $M$ images with which we must texture the target
plane. Each image has a camera matrix $P_i$ for $i=1..M$, which
translates a 3D point in the world coordinate system to a 2D point or
pixel in image $i$'s coordinates. A camera matrix $P_i$ is composed of
the camera's intrinsic parameters, such as focal length and image
center, as well as extrinsic parameters which specify the rotation and
translation of the camera's position in 3D world coordinates at the
time that image $i$ is taken. These extrinsic parameters are
determined by the backpack hardware and localization algorithms
\cite{chen2010indoor, liu2010indoor, kua2012loopclosure} and are quite
noisy.

Because the backpack system takes photos at a rate of 5 Hz, thousands
of images are available for texturing each plane. Our goal in creating
a texture mapping process is to decide which of these images should be
used, and where their contents should map onto the plane, in order to
eliminate any visual discontinuities or seams that would suggest that
the plane's final texture is not composed of a single continuous image.

\subsection{Direct Mapping}
\label{sec:directMapping}

Ignoring the fact that the camera matrices $P_{1..M}$ are inaccurate, we can texture the plane by discretizing it into small square tiles, generally about 5 pixels across, and choosing an image to texture each tile.

We choose to work with rectangular units to ensure that borders
between any two distinct images in our final texture are either
horizontal or vertical. Since most environmental features inside
buildings are horizontal or vertical, any seams in our texture
intersect them minimally and are likely to be less noticeable.

In order to select an image for texturing tile $t$, we must first
gather a list of candidate images that contain all four of its
corners, which we can quickly check by projecting $t$ into each image
using the $P_i$ camera matrices. Furthermore, each candidate image
must have been taken at a time when its camera had a clear
line-of-sight to $t$, which can be calculated using standard
ray-polygon intersection tests between the camera location, the center
of $t$, and every other plane \cite{rayintersection}.

Once we have a list of candidate images for $t$, we define a scoring
function in order to objectively select the best image. Since camera
pose errors compound over distance, we wish to minimize the distance
between cameras and the plane they texture. Additionally, we desire
images that are projected perpendicularly onto the plane, maximizing
the resolution and amount of useful texture available in their
projections, as well as minimizing any parallax effects. In other words, we wish to minimize the angle between the
plane normal and the camera axis for images selected for texture
mapping. These two criteria can be met by maximizing the function
$\frac{1}{d} (-1 \cdot \vec{c}) \cdot \vec{n}$ as shown in Figure
\ref{fig:scoringFunction}. Specifically, $d$ is the distance between
the centers of a camera and a tile, and $\vec{n}$ and $\vec{c}$ are
the directions of the plane's normal and the camera axis respectively.

\begin{figure}
  \centering
  \includegraphics[height=1.5in]{scoringFunction.jpg}
  \caption{We minimize camera angle $\alpha$ and distance $d$ by
    maximizing the scoring function $\frac{1}{d} (-1 \cdot \vec{c})
    \cdot \vec{n}$}
  \label{fig:scoringFunction}
\end{figure}

\begin{figure}[h!]
  \centering \subfloat[][]{\includegraphics[width=3in,
    height=0.9in]{wall2_naive.jpg}}

  \centering \subfloat[][]{\includegraphics[width=3in,
    height=0.9in]{wall2_cache.jpg}}

  \centering \subfloat[][]{\includegraphics[width=3in,
    height=0.9in]{wall2_cache_shift.jpg}}

  \centering \subfloat[][]{\includegraphics[width=3in,
    height=0.9in]{wall2_dynprog_noblend.jpg}}

  \centering \subfloat[][]{\includegraphics[width=3in,
    height=0.9in]{wall2_cache_shift_blend.jpg}}

  \centering \subfloat[][]{\includegraphics[width=3in,
    height=0.9in]{wall2_dynprog.jpg}}
  \caption{(a) Direct mapping. (b) Mapping with caching. (c) Mapping
    with caching after image alignment. (d) Seam minimization after
    image alignment). (e) same as (c) with blending. (f) same as (d)
    with blending.}
  \label{fig:compareAll}
\end{figure}

As Figure \ref{fig:compareAll}(a) demonstrates, this approach leads to
the best texture for each tile independently, but overall results in
many image boundaries with abrupt discontinuities, due to significant
misalignment between images, as a result of camera pose inaccuracies.

\subsection{Mapping with Caching}
\label{sec:mappingWithCaching}
Since discontinuities occur where adjacent tiles select non-aligned
images, it makes sense to take into account image selections made by
neighboring tiles while texture mapping a given tile. By using the
same image across tile boundaries, we can eliminate a discontinuity
altogether. If this is not possible because a tile is not visible in
images chosen by its neighbors, using similar images across a tile boundary will result in minimal discontinuities.

Similar to a caching mechanism, we select the best image for a tile
$t$ by searching through two subsets of images for a viable candidate,
before searching through the entire set. The first subset of images is
those selected by adjacent tiles that have already been textured. We
must first check which of these images can map to $t$, and then of
those, we make a choice according to the scoring function in Figure
\ref{fig:scoringFunction}. Before reusing this image, we ensure it
meets the criteria $\alpha < 45^\circ$, in order to be considered a
viable image, with $\alpha$ as the camera angle as shown in Figure
\ref{fig:scoringFunction}.

If no satisfactory image is found in the first subset, we check the
second subset of images, consisting of those taken near the ones in
the first subset, both spatially and temporally. These images are not
the same as the ones used for neighboring tiles, but were taken at a
similar location and time, suggesting that their localization and
projection are very similar. Again, if no viable image is found
according to the same criteria, we search the entire set of candidate
images, selecting based on the same scoring function from Figure
\ref{fig:scoringFunction}.

The result of this caching approach is shown in Figure
\ref{fig:compareAll}(b). As compared to Figure
\ref{fig:compareAll}(a), discontinuities are reduced overall, but the
amount of remaining seams suggests that image selection alone cannot
produce seamless textures. Camera matrices, or the image projections
themselves have to be adjusted in order to reliably generate clean
textures.

\section{Existing Approaches to Image Alignment}
\label{sec:existingApproaches}
In order to produce seamless texture mapping, either camera matrices
need to be refined such that their localization is pixel accurate, or
image stitching techniques need to be applied to provide this
illusion.

Before examining these approaches, we first obtain a set of images to
work with.  Rather than perform camera or image adjustments across the
many thousands of images acquired in a typical data collection, we opt
to work with the more limited set of images corresponding to those
chosen by the direct mapping approach, without caching, of Section
\ref{sec:directMapping}. This set of images constitutes a reasonable
candidate set for generating a final seamless texture since it meets
three important criteria. First, each tile on the plane is covered by
at least one image in this set; this ensures no holes in the final
texture. Second, images are all selected according to the scoring
function in Figure \ref{fig:scoringFunction}, ensuring reasonable
camera distances and angles. Third, as a side result of the scoring
function, selected images are only viable for the tiles near their
center of projection. Thus, there should be plenty of overlap between
selected images, allowing for some degree of shifting without
resulting in holes, as well as area for blending between them. With
this set of images, we now review approaches towards refining and
combining their projections.

One way to align image projections is to refine their noisy camera
poses, such that they become pixel accurate. This can be done by using
image correspondences to adjust each image's camera matrix over 6
degrees of freedom (DOF) \cite{coorg1997matching,
  liu2010indoor, shum2000systems, szeliski1997creating}. Using our same modeling datasets, recent work
followed this approach by solving a non-linear optimization problem to
refine camera poses during backpack localization \cite{liu2010indoor,
  chen2010indoor}. This approach however suffers from error
accumulation and drift as shown in Figure \ref{fig:mosaic3D}(a), as
well as high complexity and runtime.

Refining camera poses is analogous to image alignment or mosaicing,
which perform translations on images in order to align them directly.
One common approach to image alignment is to iteratively match images
on intensity differences in selected patches, often in a spatial
hierarchy \cite{hager1998efficient, lucas1981iterative,
  shum2000systems, szeliski2006image}. Another widely-used approach is
to perform feature matching, which requires the detection of visual
features and their existence in multiple images
\cite{brown2007automatic, cho2003automatic,
  mikolajczyk2005performance, szeliski2006image,
  yun2003comprehensive}. Notably, both approaches rely on the presence
of multiple visual references, as when texturing detailed objects at
high resolutions \cite{bernardinimultiplescans, wangmultipleviews}, or
when creating large scene panoramas \cite{agarwalapanoramas,
  bernardinimultiplescans, debevecviewdependent,
  snavelyphototourism}. In contrast, our indoor datasets have a high
prevalence of bare walls and ceilings, resulting in few reference
points that can be used to judge alignment.

Additionally, our datasets often contain long chains of images, which
leads to error accumulation when image correspondences are not
accurate. For example, when matching a long chain of images through
homography, a pixel in the $nth$ image must be translated into the
first image's coordinates by multiplying by the $3\times3$ matrix $H_1
H_2 H_3 ... H_n$. Any error in one of these homography matrices is
propagated to all further images, resulting in drift. Figure
\ref{fig:mosaic3D}(b) shows the output of the AutoStitch software
package, which performs homography-based image mosaicing
\cite{autostitch}. Even with many features spread across this plane,
the mosaicing produces errors that causes straight lines to appear as
waves on the plane, despite the fact that it was generated after
careful hand tuning. Many planes with fewer features simply failed
outright.

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=3in]{Graph_crop.pdf}}

  \centering \subfloat[][]{\includegraphics[width=3in]{panoMy.jpg}}
  \caption{Texture alignment via (a) the graph-based localization
    refinement algorithm from \cite{chen2010indoor} and (b) image
    mosaicing}
  \label{fig:mosaic3D}
\end{figure}

\section{Seamless Texture Mapping}
\label{sec:proposedApproach}

\begin{figure}
  \centering
  \includegraphics[width=3in]{pipeline.pdf}
  \caption{Our proposed method for seamless texture mapping.}
  \label{fig:pipeline}
\end{figure}

In this section, we describe our proposed method for seamless texture
mapping. Our approach consists of 4 steps, as seen in Figure
\ref{fig:pipeline}.  First, we choose a subset of $M$ images to work
with via the direct mapping approach of Section
\ref{sec:directMapping}. Next, we project these images onto the plane
and perform rotation and shifting in order to maximize SIFT feature
matches in overlapping areas between these projections. We then select
which textures to be used either with the tile caching method from
Section \ref{sec:mappingWithCaching}, or with the more specialized
method to be described in Section \ref{sec:seamMinimization}. This
choice depends on the availability of head-on images for a given
plane. We then finish by applying linear alpha blending to smooth out
any remaining seams.

\subsection{Image Rotation}
\label{sec:projectionAndRotation}
We begin with the projection of all images onto the plane. This is
done in the same way as the approaches in Section
\ref{sec:simpleTextureMapping}. Rather than work with all 6 DOF, which is computationally intensive, we choose to work only in two dimensions, performing 2D rotations and translations on the image projections, as errors in either lead to the greatest discontinuities.

We perform rotations using Hough transforms, which
detect the presence and orientation of linear features in our
images. Rather than match the orientation of such features in each
image, we simply apply rotations such that the strongest near-vertical
features are made completely vertical. This is effective for vertical
planes in indoor models, since they usually consist of parallel
vertical lines corresponding to doors, wall panels, rectangular
frames, etc. If features in the environment are not vertical, e.g. for
floors and ceilings, or are not parallel to each other, this step is
skipped.

\subsection{Image Shifting}
\label{sec:robustSIFTFeatureMatching}
Our next step is to align overlapping images by searching for
corresponding points between all pairs of overlapping images. We use SIFT features for their high detection rate, and choose to use feature alignment rather than intensity-based alignment due to the differences in lighting as well as possible occlusion among our images, both of which feature alignment is less sensitive to \cite{lai1999robust, lowe1999object, mikolajczyk2005performance, szeliski2006image}.

SIFT matches determine $d^x$ and $d^y$ distances between each pair of
features for two images on the plane, though these distances may not
always be the same for different features.

Since indoor environments often contain repetitive features such as
floor tiles or doors, we need to ensure that SIFT-based distances are
reliable. In order to mitigate the effect of incorrect matches and
outliers, the RANSAC framework \cite{fischler1981random} is used for a
robust estimate of the optimal $d^x_{i,j}$ and $d^y_{i,j}$ distances
between two images $i$ and $j$. The RANSAC framework handles the
consensus-building machinery, and requires a fitting function and a
distance function. For this application, the fitting function simply
finds the average distance between matches in a pair of images. The
distance function for a pair of points is chosen to be the difference
between those points' SIFT match distance and the average distance
computed by the fitting function. We use a 10 pixel outlier threshold,
so that SIFT matches are labeled as outliers if their horizontal or
vertical distances are not within 10 pixels of the average distance
computed by the fitting function.

We now use the $d^x_{i,j}$ and $d^y_{i,j}$ distances between each pair
of images to refine their positions using least squares. Recall that
there are a total of $M^{2}$ possible pairs of images, though we only
generate distances between images that overlap at SIFT feature
points. Given these distances and the original image location
estimates, we can solve a least squares problem
($\textrm{min}_{\vec{\beta}} ||A \vec{\beta} - \vec{\gamma}||_2^2 $)
to estimate the location of the images on the plane. The
$M$-dimensional vector $\vec{\beta}$ represents the unknown $x$
location of each image on the plane for $1 \dots M$. The optimal $x$
and $y$ locations are obtained in the same way, so we only consider
the $x$ locations here:

\[\vec{\beta} =
\begin{pmatrix}
  x_1, & x_2, & x_3, & \cdots & x_{M-1}, & x_M
\end{pmatrix}
\]

The $N \times (M+1)$ dimensional matrix $A$ is constructed with one
row for each pair of images with measured distances produced by the
SIFT matching stage. A row in the matrix has a $-1$ and $1$ in the
columns corresponding to the two images in the pair. For example, the
matrix below indicates that we generated a SIFT-based distance between
images 1 and 2, images 1 and 3, images 2 and 3, etc.

\[
A =
\begin{pmatrix}
  -1 & 1 & 0 & \cdots & 0 & 0\\
  -1 & 0 & 1 & \cdots & 0 & 0\\
  0 & -1 & 1 & \cdots & 0 & 0\\
  \vdots  & \vdots & \vdots & \ddots & \vdots  & \vdots\\
  0 & 0 & 0 & \cdots & 1 & 0 \\
  0 & 0 & 0 & \cdots & -1 & 1 \\
  1 & 0 & 0 & \cdots & 0 & 0 \\
\end{pmatrix}
\]

If only relative distances between images are included then there is
no way to determine the absolute location of any of the images and the
matrix becomes rank deficient. To fix this we choose the first image
to serve as the anchor for the rest, meaning all the absolute
distances are based on its original location. This is done by adding a
row with a $1$ in the first column and the rest zeros.

Finally, the $N$-dimensional observation vector $\vec{\gamma}$ is
constructed using the SIFT-based distances generated earlier in the
RANSAC matching stage. The last element in the observation vector is
the location of the first image determined by its original noisy
localization, from \cite{chen2010indoor, liu2010indoor}. Thus
$\vec{\gamma}$ can be written as:

\[
\vec{\gamma}^T =
\begin{pmatrix}
  d_{1,2}, &d_{1,3}, &d_{2,3}, &\hdots &d_{N-2,N-1}, &d_{N-1,N}, &x_1
\end{pmatrix}
\]

The $\vec{\beta}$ that minimizes $||A \vec{\beta} -
\vec{\gamma}||_2^2$ results in a set of image locations on the plane
that best honors all the SIFT-based distance measurements between
images. In practice there are often cases where there is a break in
the chain of images, meaning that no SIFT matches are found between
one segment of the plane and another. In this case we add rows to the
$A$ matrix and observations to the $\vec{\gamma}$ vector that contain
the original noisy $x$ and $y$ distance estimates generated by the
localization algorithm \cite{chen2010indoor, liu2010indoor}. Another
way to do this is to add rows for all neighboring pairs of images and
solve a weighted least squares problem where the SIFT distances are
given a higher weight i.e. 1, and the noisy distances generated by the
localization algorithm \cite{chen2010indoor, liu2010indoor} are given
a smaller weight i.e. 0.01.

After completing this same process for the $y$ dimension as well, and
making the resultant shifts, our images overlap and match each other
with far greater accuracy. Applying the tile caching method from
Section \ref{sec:mappingWithCaching} on these rotated and shifted
images results in the significant improvements shown in Figure
\ref{fig:compareAll}(c), as compared to Figure
\ref{fig:compareAll}(b).

\subsection{Seam Minimization}
\label{sec:seamMinimization}

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=1in]{projectionWall.pdf}}
  \centering
  \subfloat[][]{\includegraphics[width=1.1in]{projectionCeiling.pdf}}
  \centering
  \subfloat[][]{\includegraphics[width=0.9in]{projectionWallCrop.pdf}}
  \caption{(a) Undistorted fisheye images for vertical planes are
    tilted, but their effective camera axis is more or less normal to
    the plane. (b) The effective camera axis for ceilings is at an
    angle with respect to the plane normal. (c) Wall images are
    cropped to be rectangular.}
  \label{fig:projectionAngles}
\end{figure}

As mentioned earlier, the mobile backpack system used to capture data
in this paper contains 2 cameras with $180^\circ$ fisheye lenses
facing to the right and left. Since most human operators do not carry
the backpack in a perfectly upright position and are bent forwards at
15 to 20 degrees with respect to the vertical direction, the
undistorted fisheye images are head on with respect to vertical walls,
but at an angle with respect to horizontal floors and ceilings. This
is depicted in Figure \ref{fig:projectionAngles}. These oblique camera
angles translate into textures that span extremely large areas once
projected. Using the tile-based texture mapping criteria from Figure
\ref{fig:scoringFunction}, such projections have high or low scores
depending on the location on the plane, as shown in Figure
\ref{fig:projectionAngles}(b). The tiling approach in Section
\ref{sec:simpleTextureMapping} is thus a good choice for texturing
floors and ceilings, as it uses the parts of image projections that
are viable choices for their respective plane locations, e.g. areas
near point A in Figure \ref{fig:projectionAngles}(b), but not near
point B.

For wall planes however, most images are taken from close distances
and more or less head-on angles, resulting in higher resolution
fronto-parallel projections. As a result, for each tile on a wall
plane, the cost function of Figure \ref{fig:scoringFunction} is
relatively flat with respect to a large number of captured images, as
they are all more or less head on. Thus it is conceivable to use a
different texturing strategy for walls so as to minimize the visible
seams for texture mapped planes. One way to do so is to choose the
smallest possible set of images that (a) covers the entire plane and
(b) minimizes the amount of visible seams. A straigntforward cost
function that accomplishes the latter is the Sum of Squared pixel
Differences (SSD) in overlapping regions between all pairs of images,
after they have been aligned as described in Section
\ref{sec:projectionAndRotation}. Minimizing this cost function
encourages image boundaries to occur either in featureless areas, such
as bare walls, or in areas where images match extremely well.

To cover the entirety of a plane, our problem can be defined as
minimally covering a polygon i.e. the plane, using other polygons of
arbitrary geometry i.e. image projections, with the added constraint
of minimizing the cost function between chosen images.  This is a
complex problem, though we can take a number of steps to simplify
it. Given that wall-texture candidate images are taken from more or
less head-on angles, and assuming only minor rotations are made in
Section \ref{sec:projectionAndRotation}, we can reason that their
projections onto the plane are approximately rectangular. By cropping
them all to be rectangular, as shown in Figure
\ref{fig:projectionAngles}(c), our problem becomes the conceptually
simpler one of filling a polygon with rectangles, such that the sum of
all costs between each pair of rectangles is minimal. We thus also
retain the advantages of working with rectangular units, as explained
in Section \ref{sec:simpleTextureMapping}.

The location and orientation of the cameras and the fisheye lenses on
the acquisition backpack is such that images nearly always contain the
entirety of the floor to ceiling range of wall planes. Images are
therefore rarely projected with one above another when texturing wall
planes. In essence, we need only to ensure lateral coverage of wall
planes, e.g. from left to right, as our images provide full vertical
coverage themselves. We can thus construct a Directed Acyclic Graph
(DAG) from the images, with edge costs defined by the SSD cost
function, and solve a simple shortest path problem to find an optimal
subset of images with regard to the cost function \cite{dijkstra}.

\begin{figure}
  \centering
  \includegraphics[width=3in]{dagCreation.pdf}
  \caption{DAG construction for the image selection process.}
  \label{fig:dagCreation}
\end{figure}

Figure \ref{fig:dagCreation} demonstrates the construction of a DAG
from overlapping images of a long hallway. Images are sorted by
horizontal location left to right, and become nodes in a
graph. Directed edges are placed in the graph from left to right
between images that overlap. The weights of these edges are determined
by the SSD cost function. Next, we add two artificial nodes, one start
node representing the left border of the plane, and one end node
representing the right border of the plane. The left(right) artificial
node has directed edges with equal cost $C_0$ to(from) all images that
meet the left(right) border of the plane.

We now solve the shortest path problem from the start node to the end
node. This provides a set of images completely covering the plane
horizontally, while minimizing the cost of seams between images.

In rare cases where the vertical dimension of the plane is not
entirely covered by one or more chosen images, we are left with holes
where no images are selected to texture. To address this, we solve for
a new shortest path after modifying the edges chosen by the original
shortest path solution such that they have higher cost than all other
edges in our DAG, and solve for a new shortest path. This ensures that
these edges are not chosen again unless they are the only available
option. Our second shortest path solution results in a new set of
chosen images, of which we only retain those that cover areas not
covered by the first set. This process can be repeated as many times
as needed, until holes are no longer present. This method is not as
optimal as a 2D-coverage solution would be, but it is a fast
approximation, and adequately handles the few holes we encounter.

With this completed, we have now mapped every location on the plane to
at least one image, and have minimized the number of images, as well
as the discontinuities between their borders. As seen in Figure
\ref{fig:compareAll}(d), this seam minimization method has fewer
visible discontinuities than Figure \ref{fig:compareAll}(c)
corresponding to the tile caching approach\footnote{In Figure
  \ref{fig:compareAll}(d), we arbitrarily chose one image for
  texturing where images overlap, as blending will be discussed in the
  next section.}. This is especially evident when comparing the posters in the images, which have clear misalignment over seams in Figure \ref{fig:compareAll}(c), but are much more aligned in Figure \ref{fig:compareAll}(d). This is because the seam minimization approach
directly reduces the cost of each image boundary, while the tile
caching method uses a scoring function that only approximates this
effect. Furthermore, seam minimization guarantees the best selection
of images, while the sequential tile caching method may select images
early on that turn out to be poor choices once subsequent tiles have
been processed. The seam minimization approach is also far less
intensive in terms of memory usage and runtime, as it does not require
discretizing planes or images.

In the context of texturing an entire 3D planar model, we choose to
apply the seam minimization approach on walls, due to its superior
output when provided with head-on images. Floors and ceilings however,
given their many images taken at oblique angles, are textured using
the tile caching method.

% \begin{figure*}
%   \centering
%   \includegraphics[width=6in]{pipelineimages.pdf}
%   \caption{Textures corresponding to each step in our texture
%   mapping pipeline.}
%   \label{fig:pipelineimages}
% \end{figure*}

\subsection{Blending}
\label{sec:blending}
We now apply the same blending process to the two texturing methods:
localization refinement followed by either tile caching or seam
minimization.

Although the preprocessing steps and image selection in both methods
attempt to minimize all mismatches between images, there are
occasional unavoidable discontinuities in the final texture due to
different lighting conditions or inaccuracies in planar geometry or
projection. These can however be treated and smoothed over by applying
alpha blending over image seams.  Whether the units we are blending
are rectangularly-cropped images or rectangular tiles, we can apply
the same blending procedure, as long as we have a guaranteed overlap
between units to blend over.

For the tile caching method, we can ensure overlap by texturing a
larger tile than needed for display. For example, for a rendered tile
$l_1 \times l_1$, we can associate it with a texture $(l_1 + l_2)
\times (l_1 + l_2)$ in size. For the seam minimization method, we have
already ensured overlap between images. To enforce consistent blending
however, we add a minimuum required overlap distance while solving the
shortest path problem in Section
\ref{sec:seamMinimization}. Additionally, if images overlap in a
region greater than the overlap distance, we only apply blending over
an area equal to the overlap distance.

After blending pixels linearly across overlapping regions using alpha
blending, texture mapping is complete. Figures \ref{fig:compareAll}(e)
and \ref{fig:compareAll}(f) show the blended versions of Figures
\ref{fig:compareAll}(c) and \ref{fig:compareAll}(d) respectively. It
is clear that the seam minimization approach still exhibits better
alignment and fewer seams than the tile-caching method, as Figure
\ref{fig:compareAll}(f) has the best visual quality among the textures
in Figure \ref{fig:compareAll}.

\section{Results and Conclusions}
\label{sec:resultsAndConclusions}
Examples of ceilings and floors textured with the tile caching
approach, and walls textured with the seam minimization approach, are
displayed in Figure \ref{fig:results}. High resolution texture
comparisons, as well as a walkthrough of a fully textured 3D model are
available in the accompanying video to this paper.

In this paper, we have developed an approach to texture map models
with noisy camera localization data. We are able to refine image
locations based on feature matching, and robustly handle outliers. The
tile-based mapping approach can be used to texture both simple
rectangular walls as well as complex floor and ceiling geometry. We
also presented a seam minimization texturing method that produces
seamless textures on vertical planes where multiple head-on images are
available. Each of these approaches is highly modular, and easily
tunable for different environments and acquisition hardware.

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=3in]{finalfloors.jpg}}

  \centering \subfloat[][]{\includegraphics[width=3in]{4thfloor8.jpg}}

  \centering
  \subfloat[][]{\includegraphics[width=3in]{floorcropped.jpg}}

  \centering \subfloat[][]{\includegraphics[width=3in]{fullmodel.png}}
  \caption{Examples of our final texture mapping output for (a) walls,
    (b) a ceiling, (c) an entire floor, (d) a full model}
  \label{fig:results}
\end{figure}

{\small \bibliographystyle{ieee} \bibliography{egbib} }

\appendix
\section*{Appendix A. Occlusion Masking}
\label{sec:occlusionMasking}
For the seam minimization process, it is important that images contain
only content that should be mapped onto the target plane in
question. The tiling approach checks occlusion for each tile as it is
being textured, whereas the seam minimization approach uses entire
images.  Occlusion checks thus need to be performed over the entirety
of each image to determine available areas for texture mapping.

Fortunately, by virtue of our indoor environments, the vast majority
of surface geometry is either horizontal or vertical, with high
amounts of right angles. This means that after masking out occluded
areas, image projections will remain largely rectangular. We can be
efficient by recursively splitting each image into rectangular pieces,
and performing the same occlusion checks used in the tiling process
where needed. To occlude out rectangular pieces, their texture is
removed, as untextured areas are never used for texture mapping.

\end{document}