### 2012-12-30 spie/paper.tex
few more holes to fill

%  article.tex (Version 3.3, released 19 January 2008)
%  Article to demonstrate format for SPIE Proceedings
%  Special instructions are included in this file after the
%  symbol %>>>>
%  Numerous commands are commented out, but included to show how
%  to effect various options, e.g., to print page numbers, etc.
%  This LaTeX source file is composed for LaTeX2e.

%  The following commands have been added in the SPIE class 
%  file (spie.cls) and will not be understood in other classes:
%  \supit{}, \authorinfo{}, \skiplinehalf, \keywords{}
%  The bibliography style file is called spiebib.bst, 
%  which replaces the standard style unstr.bst.

\documentclass[]{spie}  %>>> use for US letter paper
%%\documentclass[a4paper]{spie}  %>>> use this instead for A4 paper
%%\documentclass[nocompress]{spie}  %>>> to avoid compression of citations
%% \addtolength{\voffset}{9mm}   %>>> moves text field down
%% \renewcommand{\baselinestretch}{1.65}   %>>> 1.65 for double spacing, 1.25 for 1.5 spacing 
%  The following command loads a graphics package to include images 
%  in the document. It may be necessary to specify a DVI driver option,
%  e.g., [dvips], but that may be inappropriate for some LaTeX 
%  installations. 
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}

\title{Texture mapping 3D planar models of indoor environments with noisy camera poses}

%>>>> The author is responsible for formatting the 
%  author list and their institutions.  Use  \skiplinehalf 
%  to separate author list from addresses and between each address.
%  The correspondence between each author and his/her address
%  can be indicated with a superscript in italics, 
%  which is easily obtained with \supit{}.

\author{Peter Cheng, Michael Anderson, Stewart He, Avideh Zakhor
\skiplinehalf
University of California, Berkeley\\
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
%>>>> uncomment following for page numbers
% \pagestyle{plain}    
%>>>> uncomment following to start page numbering at 301 
%\setcounter{page}{301}

\begin{document}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{abstract}

  Automated 3D modeling of building interiors is used in applications
  such as virtual reality and environment mapping. Texturing these
  models allows for photorealistic visualizations of the data
  collected by such modeling systems. Camera poses obtained by these
  systems often suffer from inaccuracies, resulting in visible
  discontinuities when successive images are projected adjacently onto
  a surface for texturing. Existing methods to stitch images together
  are often computationally expensive and work independently of pose
  estimates and geometry information. We propose a method to refine
  camera poses using both existing estimates and geometry information,
  followed by two different methods to composite images together,
  based on the uniformity of available images. The effectiveness of
  our methods is demonstrated on a number of different indoor
  environments.
\end{abstract}

% >>>> Include a list of keywords after the abstract

\keywords{Texture Mapping, Reconstruction, Image Stitching, Mosaicing}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:introduction} % \label{} allows reference to this section

Three-dimensional modeling of indoor environments has a variety of
applications such as training and simulation for disaster management,
virtual heritage conservation, and mapping of hazardous sites. Manual
construction of these digital models can be time-consuming, and as a
result, automated 3D site modeling has garnered much interest in
recent years.

The first step in automated 3D modeling is the physical scanning of an
environment's geometry. An indoor modeling system must recover its own
poses within an environment while simultaneously reconstructing the 3D
structure of the environment itself. \cite{chen2010indoor,
  liu2010indoor, kua2012loopclosure}. This is known as the
simultaneous localization and mapping (SLAM) problem, and is generally
solved by taking readings from laser range scanners, cameras, and
inertial measurement units (IMUs) at multiple locations within the
environment.

In this paper, we work with data obtained from a backpack-mounted
system, carried by an ambulatory human. Such a system provides
advantages over more common wheel-mounted systems in terms of agility
and portability, but results in much higher localization error
\cite{liu2010indoor}. As a result, common methods for texture mapping
generally produce poor results, leading to the development of the
approaches contained in this paper. Besides localization information
and images, environment geometry is required as well for texture
mapping. In this paper we work mainly with models obtained by fitting
planar surfaces to point clouds generated by the backpack system
\cite{sanchez2012point}. We have also developed an appropriate method
to work with near-planar triangular meshes, as described in Appendix
A.

Though our texture mapping procedure was designed with the
aforementioned system in mind, it is applicable to other systems,
provided the required inputs of planar geometry, images with rough
extrinsic pose estimates, and either known or estimated intrinsic
camera calibration matrices. An important benefit of our method is its
relatively low complexity and runtime, and independence of many
steps, allowing for quick tuning of parameters and easy
accomodation of new input data given updates in localization or
model-generation algorithms.

The remainder of the paper is organized as follows. Section
\ref{sec:simpleTextureMapping} explains how input images are projected
onto our geometry and simple approaches towards texturing. Section
\ref{sec:existingApproaches} covers more sophisticated existing
approaches to image matching and stitching, and their performance on
our datasets. Section \ref{sec:imageAlignment} contains our approach
towards image alignment, followed by Section
\ref{sec:imageCompositing}, which describes compositing
approaches. Section \ref{sec:resultsAndConclusions} contains further
results and conclusions, as well as areas of ongoing and future work.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Simple Texture Mapping}
\label{sec:simpleTextureMapping}

The geometry of the texture mapping process for a plane is shown in
Figure \ref{fig:projection}.  As described earlier, we are provided
with a set of $M$ images to texture the target plane. Each image has a
camera matrix $P_i$ for $i=1..M$, which translates a 3D point in the
world coordinate system to a 2D point or pixel in image $i$'s
coordinates. A camera matrix $P_i$ is composed of the camera's
intrinsic parameters, such as focal length and image center, as well
as extrinsic parameters which specify the rotation and translation of
the camera's position in 3D world coordinates at the time that image
$i$ is taken. These extrinsic parameters are determined by the
backpack hardware and localization algorithms \cite{chen2010indoor,
  liu2010indoor, kua2012loopclosure} and are quite noisy.

Because our backpack system takes photos at a rate of 5 Hz, thousands
of images are available for texturing each surface in our model. Our
goal in designing a texture mapping process is to decide which of
these images should be used, and where their contents should map onto
the texture, in order to minimize any visual discontinuities or seams
that would suggest that the plane's final texture is not composed of a
single continuous image.

\begin{figure}
  \begin{minipage}[b]{0.45\linewidth}
  \centering
  \includegraphics[height=2in]{Projection.pdf}
  \caption{Planes are specified in 3D space by four corners $C_1$ to
    $C_4$. Images are related to each plane through the camera matrics
    $P_{1..m}$. }
  \label{fig:projection}
  \end{minipage}
  \hspace{0.5cm}
  \begin{minipage}[b]{0.45\linewidth}
  \centering
  \includegraphics[height=1.5in]{scoringFunction.jpg}
  \caption{We minimize camera angle $\alpha$ and distance $d$ by
    maximizing the scoring function $\frac{1}{d} (-1 \cdot \vec{c})
    \cdot \vec{n}$}
  \label{fig:scoringFunction}
  \end{minipage}
\end{figure}

\subsection{Tile-based Texture Mapping}
\label{sec:tileBasedMapping}
Ignoring the fact that the camera matrices $P_{1..M}$ are inaccurate,
we can texture the plane by discretizing it into small square tiles,
in our case 5 pixels across, and choosing an image to texture each
tile.

We choose to work with rectangular units to ensure that borders
between any two distinct images in the final texture are either
horizontal or vertical. Since most strong environmental features
inside buildings are horizontal or vertical, any visible seams in our
texture intersect them minimally and are less noticeable.

In order to select an image for texturing a tile $t$, we must first
gather a list of candidate images that contain all four of its
corners, which we can quickly check by projecting $t$ into each image
using the $P_i$ camera matrices. Furthermore, each candidate image
must have been taken at a time when its camera had a clear
line-of-sight to $t$, which can be calculated using standard
ray-polygon intersection tests between the camera location, $t$, and
every other plane \cite{rayintersection}.

Once we have a list of candidate images for $t$, we define a scoring
function in order to objectively select the best image for texturing
$t$. Since camera pose errors compound over distance, we wish to
minimize the distance between cameras and the surfaces they
texture. Additionally, we desire images that are projected
perpendicularly onto the plane, maximizing the resolution and amount
of useful texture available in their projections, as well as
minimizing any parallax effects due to real-world geometry not
accurately represented by our digital model. In other words, we wish
to minimize the angle between the tile's normal vector and the camera
axis for images selected for texture mapping. These two criteria can
be met by maximizing the function $\frac{1}{d} (-1 \cdot \vec{c})
\cdot \vec{n}$ as shown in Figure
\ref{fig:scoringFunction}. Specifically, $d$ is the distance between
the centers of a camera and a tile, and $\vec{n}$ and $\vec{c}$ are
the directions of the plane's normal and the camera axis respectively.

\begin{figure}
  \centering 
  \subfloat[][]{\includegraphics[width=3in, height=0.97in]{wall2_naive.pdf}}
  \subfloat[][]{\includegraphics[width=3in, height=0.97in]{wall2_naive_shift.pdf}}

  \centering
  \subfloat[][]{\includegraphics[width=3in, height=0.97in]{wall2_cache_shift.pdf}}
  \subfloat[][]{\includegraphics[width=3in, height=0.97in]{wall2_shortest_shift.pdf}}

  \centering 
  \subfloat[][]{\includegraphics[width=3in, height=0.97in]{wall2_cache_shift_blend.pdf}}
  \subfloat[][]{\includegraphics[width=3in, height=0.97in]{wall2_shortest_shift_blend.pdf}}
  \caption{(a) Tile-based texturing. (b) Tile-based texturing after
    image alignment. (c) Tile-based texturing after image alignment
    with caching. (d) Shortest path texturing after image
    alignment). (e) Same as (c) with blending. (f) Same as (d) with
    blending.}
  \label{fig:compareAll}
\end{figure}

As Figure \ref{fig:compareAll}(a) demonstrates, this approach leads to
the best texture for each tile independently, but results in many
image boundaries with abrupt discontinuities between tiles, due to
significant misalignment between images. Given our high number of
input images, it makese sense to further refine the image selection
procedure, as performed in Section \ref{sec:imageCompositing}, but for
our solution to be more robust, we first perform image alignment in
order maximize the accuracy of each image.

\section{Existing Approaches to Image Alignment}
\label{sec:existingApproaches}
Stitching together multiple images to produce a larger image is a
commonly performed task, with many successful approaches over the past
few decades. Parts of images are first matched to eachother, through
direct pixel comparisons, or more commonly through feature detection
and matching. Images are then adjusted to maximize matches, either by
calculating homographies between pairs of images, or by modifying
camera poses in 1 to 6 degrees of freedom.

Feature detection and matching works best when multiple unique visual
references exist in the environment and are present within multiple
images. In contrast, our indoor environments have a high prevalence of
bare surfaces, as well as repeating textures that cause difficulty in
disambiguating features. This lack of strong reference points results
in high uncertainty when matching images together.

Additionally, our datasets often contain long chains of images, which
leads to error accumulation when image correspondences are not
accurate. For example, when matching a long chain of images through
homographies, a pixel in the $nth$ image must be translated into the
first image's coordinates by multiplying by the $3\times3$ matrix $H_1
H_2 H_3 ... H_n$. Any error in one of these homography matrices is
propagated to all further images, resulting in drift. Figure
\ref{fig:mosaic3D}(b) shows the output of the AutoStitch software
package, which performs homography-based image mosaicing
\cite{autostitch}. Even with many features spread across this surface,
the mosaicing produces errors that cause straight lines to appear
distorted, despite the fact that it was generated after careful hand
tuning. Many areas with fewer visual features simply failed outright
using this approach.

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=3in]{Graph_crop.pdf}}
  \subfloat[][]{\includegraphics[width=3in]{panoMy.jpg}}

  \caption{Texture alignment via (a) the graph-based localization
    refinement algorithm from \cite{chen2010indoor} and (b) image
    mosaicing.}
  \label{fig:mosaic3D}
\end{figure}

Image projections can also be aligned by iteratively adjusting camera
poses to maximize matches. This is generally done over 6 degrees of
freedom \cite{liu2010indoor}. The approach in the cited paper works
well locally, given high feature density, but over larger areas
results again in error accumulation and drift, especially in
feature-sparse areas, as shown in Figure
\ref{fig:mosaic3D}(a). Furthermore, it adjusts image locations using
an iterative process, which results in high runtime.

\section{2D Image Alignment with Geometry Information}
\label{sec:2dAlignment}
In this section, we describe our method of efficient and robust image
alignment. Our approach consists of three main parts. First, all
images are projected onto the surface and lines within these
projections are detected. These lines are then matched up with
geometry-based lines comprising the surface's boundary and
intersection with other surfaces. Second, occlusion checks are
performed to remove invalid parts of each image for the target
surface. These two steps are image-independent and performed in
parallel. Third, we detect SIFT feature matches between pairs of
images and solve a linear least squares problem in 2D to maximize
matches.

All of our following calculations and alignments are performed in 2D,
partly for efficiency, and partly because the nature of our input data
is such that localization error chiefly occurs in two dimensions, in
the same plane as the surface being projected onto. Recall that our
data acquisition system is backpack-mounted, with cameras facing to
the sides of the operator. The operator makes efforts to walk parallel
to walls, and the localization and model-generation algorithms that
provide our input often conform the operator's path to be straight and
parallel to detected surfaces \cite{kua2012loopclosure,
  sanchez2012point}. Since the operator walks in as upright a position
as possible, errors in roll, yaw, and the operator's distance from
parallel surfaces are minimal. Thus, our highest errors stem from
uncertainty in the operator's pitch, equivalent to rotation around the
camera axes, as well as location along surfaces. These equate to 2D
rotation and translation in a surface's plane.

\subsection{Geometry-based Alignment}
\label{sec:geometryAlignment}
After computing each image's projection onto the target surface, as
described in Section \ref{sec:simpleTextureMapping}, we detect lines
in the projections using Hough transforms. Experience and intuition
show that walls in indoor environments contain linear features that
are either horizontal or vertical, often corresponding to doors,
windows, posters, etc. Thus, when texturing walls, we rotate images in
2D such that dominant lines are made to be horizontal or
vertical. This can be further extrapolated by orienting lines with a
wall's boundaries, for example in areas with a slanted roof.

\begin{figure}
  \centering
  \includegraphics[height=1.5in]{geometryFeature.jpg}
  \includegraphics[height=1.5in]{imageCorner.jpg}
  \includegraphics[height=1.5in]{geometryImageAlign.jpg}
  \caption{(a) The red lines indidcate geometry-based lines,
    consisting of boundaries and intersections. (b) A matching
    image-based line, corresponding to a corner. (c) The image rotated
    and shifted such that the matched lines align align with
    geometry.}
  \label{fig:geometryAlignment}
\end{figure}

At this point in time, image occlusions have not been accounted
for. As a result, some image projections contain texture that should
project to an adjacent surface, generally with a linear boundary where
the two surfaces meet. If this linear boundary is detected by Hough
transform, the image is rotated and shifted in 2D such that the visual
boundary between two surfaces matches the physical boundary in our
digital model. An example of such an adjustment is in Figure
\ref{fig:geometryAlignment}

To perform the rotation and alignment, one method would be to detect
all line matches and solve for an optimal homography. In nearly all
cases however, we find that a single 2D rotation is sufficient to
properly align lines in an image. We also rarely have more than one
linear geometry feature to match to in each orthogonal
direction. Thus, for a rotation angle, we simply take all angles
between image-based and geometry-based lines under 5 degrees, and
average them. After the image is rotated, we find the horizontal and
vertical image-based and geometry-based line segments with minimum
orthogonal distance between them. For both distances, if it is under 5
mm, we apply the corresponding shift, and indicate that the image was
aligned to geometry in one or both dimensions, which reduces further
shifting in Section \ref{sec:robustSIFTFeatureMatching}.

\subsection{Image Occlusion}
\label{sec:imageOcclusion}
Now that images have been aligned to geometry where possible, we
perform a simple recursive occlusion procedure on each image. This is
done by performing the intersection tests in section
\ref{sec:simpleTextureMapping} in a regularly spaced grid. Where four
corners of a rectangular region are occluded, texture is
removed. Where no corners are occluded, nothing occurs. Where there is
a mixture of both, the rectangular region is subdivided into four, and
the same process is performed on each. By performing Section
\ref{sec:geometryAlignment}'s alignment procedure before occlusion,
texture belonging to other surfaces is accurately removed, which is
necessary for the next section.

\subsection{2D Feature Alignment}
\label{sec:robustSIFTFeatureMatching}
Our next step is to align overlapping images by searching for
corresponding points between all pairs of overlapping images. We use
SIFT features for their high detection rate, and choose to use feature
alignment rather than pixel or intensity-based alignment due to the
differences in lighting as well as possible occlusion among our
images, both of which feature alignment is less sensitive to
\cite{lowe1999object, mikolajczyk2005performance, szeliski2006image}.
We use SiftGPU \cite{siftgpu} for its high performance on both feature
detection as well as pairwise matching. These matches determine $d^x$
and $d^y$ distances between each pair of features for two image
projections, though these distances may not always be the same for
different features. Since indoor environments often contain repetitive
features such as floor tiles or doors, we need to ensure that
SIFT-based distances are reliable. First, we only perform alignment on
pairs of images that overlap given the original noisy poses. Second,
we discard feature matches that correspond to an image distance
greater than 40 pixels from what the noisy poses estimate. In order to
utilize the remaining feature matches robustly, the RANSAC framework
\cite{fischler1981random} is used to estimate the optimal $d^x_{i,j}$
and $d^y_{i,j}$ distances between two images $i$ and $j$. The RANSAC
framework requires a fitting function and a distance function. For
this application, the fitting function finds the average distance
between matches in a pair of images, and the distance function for a
pair of points is chosen to be the difference between those points'
SIFT match distance and the average distance computed by the fitting
function. We use a 10 pixel threshold, so that SIFT matches are
labeled as outliers if their horizontal or vertical distances are not
within 10 pixels of the average distance computed by the fitting
function.

We now use the calculated $d^x_{i,j}$ and $d^y_{i,j}$ distances
between each pair of images to refine their positions using weighted
linear least squares. There are a total of $M^{2}$ possible pairs of
images, though we only generate distances between images with SIFT
matches. Given these distances and the original image location
estimates, we can solve a least squares problem
($\textrm{min}_{\vec{\beta}} ||A \vec{\beta} - \vec{\gamma}||_2^2 $)
to estimate the location of the images on the plane. The
$M$-dimensional vector $\vec{\beta}$ represents the unknown $x$
location of each image on the plane for $1 \dots M$. The optimal $x$
and $y$ locations are obtained in the same way, so we only consider
the $x$ locations here:

\[\vec{\beta} =
\begin{pmatrix}
  x_1, & x_2, & x_3, & \cdots & x_{M-1}, & x_M
\end{pmatrix}
\]

The $N \times M$ matrix $A$ is constructed with one row for each pair
of images with measured distances produced by the SIFT matching
stage. A row in the matrix has a $-1$ and $1$ in the columns
corresponding to the two images in the pair. For example, the matrix
below indicates a SIFT-based distance between images 1 and 2, images 1
and 3, images 2 and 3, etc.
\[
A =
\begin{pmatrix}
  -1 & 1 & 0 & \cdots & 0 & 0\\
  -1 & 0 & 1 & \cdots & 0 & 0\\
  0 & -1 & 1 & \cdots & 0 & 0\\
  \vdots  & \vdots & \vdots & \ddots & \vdots  & \vdots\\
  0 & 0 & 0 & \cdots & 1 & 0 \\
  0 & 0 & 0 & \cdots & -1 & 1 \\
  1 & 0 & 0 & \cdots & 0 & 0\\
\end{pmatrix}
\]
If only relative distances between images are included, the absolute
location of the images can not be calculated, and the matrix is rank
deficient. From Section \ref{geometryBasedAlignemtn}, a number of
images are anchored to geometry points, in one or both dimensions, and
thus their locations can be used to fix the rest in place. In case no
anchor images exist, we simply arbitrarily pick an image, as we have
no other reference points to work with. In the above matrix, the first
image is set to be such an anchor, simply by placing a $1$ in its
column.

The $N$-dimensional observation vector $\vec{\gamma}$ is constructed
using the SIFT-based distances generated in the RANSAC matching
stage. Elements in the observation vector corresponding to anchor
images are simply their locations as determined by the original noisy
localization. Thus a $\vec{\gamma}$ corresponding to the above matrix
can be written as:

\[
\vec{\gamma}^T =
\begin{pmatrix}
  d_{1,2}, &d_{1,3}, &d_{2,3}, &\hdots &d_{N-2,N-1}, &d_{N-1,N}, &x_1
\end{pmatrix}
\]

The $\vec{\beta}$ that minimizes $||A \vec{\beta} -
\vec{\gamma}||_2^2$ results in a set of image locations on the plane
that best honors all the SIFT-based distance measurements between
images. This solution however does not make use of our noisy camera
poses, and will fail when no SIFT matches are found between one
segment of the plane and another. To account for this, we add rows to
the $A$ matrix and observations to the $\vec{\gamma}$ vector
corresponding to the original noisy distances. We then add weighting
to our least squares problem where the SIFT distances and anchor
values are given a high weight e.g. 1, while the noisy distances are
given a smaller weight e.g. 0.01.

After completing this same process for the $y$ dimension as well, and
making the resultant shifts, our images overlap and match each other
with far greater accuracy. Applying the simple mapping scheme in
Section \ref{sec:tileBasedMapping} to the same wall used in that section results in Figure \ref{fig:compareAll}(b), which has far
fewer seams, though some are still visible.

\section{Image Compositing}
\label{sec:imageCompositing}
In this section, we revisit the tile-based texturing approach from
Section \ref{sec:simpleTextureMapping}, with an added caching
mechanism to reduce image boundaries. This method works well given all
manner of camera poses and surfaces, but for optimal cases where we
have large sections of usable texture from images, we propose a
superior method that further reduces image boundaries.

\subsection{Tile-Mapping with Caching}
\label{sec:mappingWithCaching}
From Section \ref{sec:tileBasedMapping}, we saw that discontinuities
occur where adjacent tiles are textured by different images. Though
Section \ref{sec:2dAlignment}'s image alignment removes many such
discontinuities, there are still cases where seams are visible due to
imprecise matching or other factors such as model-based errors. To
reduce the cases where this happens, it makes sense to take into
account image selections made by neighboring tiles while texture
mapping a given tile. By using the same image across tile boundaries,
we can eliminate a discontinuity altogether. If this is not possible
because a tile is not visible in images used by neighboring tiles,
using similar images across tile boundaries also leads to less
noticeable discontinuities.

Essentially a caching mechanism, we select the best image for a tile
$t$ by searching through two subsets of images for a viable candidate,
before searching through the entire set of valid images. The first
subset of images is the images selected by adjacent tiles that have
already been textured. We must first check which of these images can
map to $t$, and then of those, we make a choice according to the
scoring function in Figure \ref{fig:scoringFunction}. Before reusing
this image, we ensure it meets the criteria $\alpha < 45^\circ$, in
order to ensure a high resolution projection, with $\alpha$ as the
camera angle as shown in Figure \ref{fig:scoringFunction}.

If no satisfactory image is found in the first subset, we check a
second subset of images, consisting of those taken near the ones in
the first subset, both spatially and temporally. These images are not
the same as the ones used for neighboring tiles, but are taken at a
similar location and time, suggesting that their localization and
projection are quite similar, and thus likely matched more cleanly. If
no viable image is found according to the same criteria as before, we
search the entire set of candidate images, selecting based on the same
scoring function from Figure \ref{fig:scoringFunction}.

The result of this caching approach is shown in Figure
\ref{fig:compareAll}(c), where seams are now greatly reduced as
compared to Figure \ref{fig:compareAll}(b).

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=1in]{projectionWall.pdf}}
  \centering
  \subfloat[][]{\includegraphics[width=1.1in]{projectionCeiling.pdf}}
  \centering
  \subfloat[][]{\includegraphics[width=0.9in]{projectionWallCrop.pdf}}
  \caption{(a) Images for vertical planes are tilted, but their camera
    axes are more or less normal to their respective planes. (b)
    Camera axes for ceiling images are at large angles with respect to
    plane normals. (c) Wall images are cropped to be rectangular.}
  \label{fig:projectionAngles}
\end{figure}

As mentioned earlier, our data comes from a mobile backpack
system. Human operators can not carry the backpack in a perfectly
upright position and are bent forwards at 15 to 20 degrees with
respect to the vertical direction. As a result, cameras facing
sideways are head on with respect to vertical walls, while cameras
oriented towards the top or bottom of the backpack are at an angle
with respect to horizontal floors and ceilings. This is depicted in
Figures \ref{fig:projectionAngles}(a) and
\ref{fig:projectionAngles}(b). These oblique camera angles for
horizontal surfaces translate into textures that span large areas once
projected, as shown in Figure \ref{fig:projectionAngles}(b). Using the
tile-based texture mapping criteria from Figure
\ref{fig:scoringFunction}, such projections have highly varying scores
depending on the location of a tile on the plane. Thus, the tiling
approach in this section is a good choice for texturing floors and
ceilings, as it uses the parts of image projections that maximize
resolution and accuracy for their respective plane locations,
e.g. areas near point A and not near point B, in Figure \ref{fig:projectionAngles}(b).

\subsection{Shortest Path Texturing}
\label{sec:shortestPath}
For vertical walls, most images are taken from close distances and
head-on angles, resulting in high resolution fronto-parallel
projections. As a result, for each tile on a wall plane, the scoring
function of Figure \ref{fig:scoringFunction} is relatively flat with
respect to candidate images, as they are all more or less head
on. Thus, the scoring function is less significant for walls, and it
is conceivable to use a different texturing strategy to directly
minimize visible seams when texturing them. This is done by choosing
the smallest possible set of images that (a) covers the entire plane
and (b) minimizes the visibility of borders between them. A
straightforward cost function that accomplishes the latter is the sum
of squared differences (SSD) of pixels in overlapping regions between
all pairs of images. Minimizing this cost function encourages image
boundaries to occur either in featureless areas, such as bare walls,
or in areas where images match extremely well.

The first step in this procedure is to obtain a list of images useful
for texturing the wall. A simple way to do this is to use the images
selected by the tiling process in Section
\ref{sec:tileBasedMapping}. Such a list of images is guaranteed to
cover the entire wall, and consists of desired camera poses overall.

To cover the entirety of a plane, our problem can be defined as
minimally covering a polygon i.e. the planar surface, using other
polygons of arbitrary geometry i.e. image projections, with the added
constraint of minimizing the cost function between chosen images.
This is a complex problem, though we can take a number of steps to
simplify it.

Given that wall-texture candidate images are taken from more or less
head-on angles, and knowing that only minor rotations are made in
Section \ref{sec:2dAlignment}, we can reason that their projections
onto the plane are approximately rectangular. By cropping them all to
be rectangular, as shown in Figure \ref{fig:projectionAngles}(c), our
problem becomes the conceptually simpler one of filling a polygon with
rectangles, such that the sum of all costs between each pair of
rectangles is minimal. We thus also retain the advantages of working
with rectangular units, as explained in Section
\ref{sec:tileBasedMapping}.

The operator's path, and the location and orientation of the cameras
on the acquisition backpack is such that images nearly always contain
the entirety of the floor to ceiling range of wall planes. Images
therefore rarely need to be projected with one above another when
texturing wall planes. In essence, we need only to ensure lateral
coverage of wall planes, e.g. from left to right, as our images
provide full vertical coverage themselves. We can thus construct a
Directed Acyclic Graph (DAG) from the images, with edge costs defined
by the SSD cost function, and solve a simple shortest path problem to
find an optimal subset of images with regard to the SSD cost function
\cite{dijkstra}.

\begin{figure}
  \centering
  \includegraphics[width=3in]{dagCreation.pdf}
  \caption{DAG construction for the image selection process. \\}
  \label{fig:dagCreation}
\end{figure}

Figure \ref{fig:dagCreation} demonstrates the construction of a DAG
from overlapping images of a hallway wall. Images are sorted by
horizontal location left to right, and become nodes in a
graph. Directed edges are placed in the graph from left to right
between overlapping images. The weights of these edges are determined
by the SSD cost function. Next, we add two artificial nodes, one start
node representing the left border of the plane, and one end node
representing the right border of the plane. The left(right) artificial
node has directed edges with equal and arbitrary cost $C_0$ to(from)
all images that meet the left(right) border of the plane.

We now solve the shortest path problem from the start node to the end
node. This results in a set of images completely covering the plane
horizontally, while minimizing the cost of seams between images.

In rare cases where the vertical dimension of the plane is not
entirely covered by one or more chosen images, we are left with holes
where no images are selected to texture. Since these holes are rare,
and generally fairly small, we use a greedy approach, repeatedly
filling the hole with images that result in the lowest SSD costs in a
blending region around the hole, as discussed in Section
\ref{sec:blending}. This method is not as optimal as a true
2D-coverage solution would be, but it is a fast approximation, and
adequately handles the few holes we encounter.

With this completed, we have now mapped every location on the plane to
at least one image, and have minimized the number of images, as well
as the discontinuities at their borders. As seen in Figure
\ref{fig:compareAll}(d), this shortest path method has fewer visible
discontinuities than Figure \ref{fig:compareAll}(c) corresponding to
the tile caching approach\footnote{In Figure \ref{fig:compareAll}(d),
  we arbitrarily chose one image for texturing where images overlap,
  as blending will be discussed in section \ref{sec:blending}.}. This
is especially evident when comparing the posters in the images, which
have clear misalignment over seams in Figure \ref{fig:compareAll}(c),
but are much more aligned in Figure \ref{fig:compareAll}(d). This
shortest path approach approach directly reduces the cost of each
image boundary, while the tile caching method uses a scoring function
that only approximates this effect. Furthermore, this approach
guarantees the best selection of images to minimize seams, while the
sequential tile caching method may select images early on that turn
out to be poor choices once subsequent tiles have been processed. This
shortest path approach is also far less intensive in terms of memory
usage and runtime, both during texture generation and rendering, as it
does not require discretizing planes or images.

When texturing an entire 3D planar model, we apply the shortest path
method on walls, due to its superior output when provided with head-on
images. Floors and ceilings however, given their many images taken at
oblique angles, are textured using the tile caching method.

\subsection{Blending}
\label{sec:blending}
We now apply a blending procedure to both texturing methods. Although
the image alignment steps and image selection in both methods attempt
to minimize all mismatches between images, there are occasional
unavoidable discontinuities in the final texture due to different
lighting conditions or inaccuracies in model geometry. These can
however be treated and smoothed over by applying alpha blending over
image seams.  Whether the units we are blending are
rectangularly-cropped images or rectangular tiles, we can apply the
same blending procedure, as long as we have a guaranteed overlap
between units to blend over.

For the tile caching method, we can ensure overlap by texturing a
larger tile than needed for display. For example, for a rendered tile
$l_1 \times l_1$, we can associate it with a texture $(l_1 + l_2)
\times (l_1 + l_2)$ in size.  We have found $l_2 = \frac{l_1}{2}$ to
provide a balance between blending and keeping features unblurred. For
the shortest path method, we have already ensured overlap between
images. To enforce consistent blending however, we add a minimum
required overlap distance of 40 px while solving the shortest path
problem in Section \ref{sec:shortestPath}. Additionally, if images
overlap in a region greater than the overlap distance, we only apply
blending over an area equal to the overlap distance.

After performing linear alpha blending across overlapping regions, our
texture mapping process is complete. Figures \ref{fig:compareAll}(e)
and \ref{fig:compareAll}(f) show the blended versions of Figures
\ref{fig:compareAll}(c) and \ref{fig:compareAll}(d) respectively. It
is clear that the shortest path approach exhibits better alignment and
fewer seams than the tile-caching method, as Figure
\ref{fig:compareAll}(f) has the best visual quality among the textures
in Figure \ref{fig:compareAll}.

\section{Results and Conclusions}
\label{sec:resultsAndConclusions}
Examples of ceilings and floors textured with the tile caching
approach, and walls textured with the shortest path approach, are
displayed in Figure \ref{fig:results}. High resolution texture
comparisons, as well as a walkthrough of a fully textured 3D model are
available in the accompanying video to this paper.

As mentioned earlier, our approach is fairly efficient. The top wall
in Figure \ref{fig:results}(a) is 7543 $\times$ 776 pixels, and spans a
40-meter long wall. Given 41000 input images, it takes approximately a
minute to pick 36 candidate images, followed, by another minute to
perform the shortest path texturing, both on a consumer-grade 2.8GHz
dual-core laptop with 4GB RAM. While not fast enough for real-time
visualization, the procedure is quick enough for quick evaluation
given changes in various internal parameters or modifications to input
data.

\subsection{Future stuff}
For the sake of simplicity, - cases where multiple whatever are few,
but for true generality, should go to 3d adjustments

\begin{figure*}
  \centering
  \subfloat[][]{\includegraphics[width=3in]{finalfloors.jpg}} ~~~~~~~~
  \centering
  \subfloat[][]{\includegraphics[width=3in]{finalceilings.jpg}}

  \centering \subfloat[][]{\includegraphics[height=2in,
    width=3in]{floorcropped.jpg}} ~~~~~~~~ \centering
  \subfloat[][]{\includegraphics[width=3in]{fullmodel.png}}
  \caption{Examples of our final texture mapping output for (a) walls,
    (b) ceilings, (c) an entire floor, (d) a full model.}
  \label{fig:results}
\end{figure*}

\appendix
\section{Working with Near-Planar Geometry}
do stuff and stuff

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%% References %%%%%

\bibliography{report} %>>>> bibliography data in report.bib
\bibliographystyle{spiebib} %>>>> makes bibtex use spiebib.bst

\end{document}