### 2013-01-18 spie/newpaper.tex

%  article.tex (Version 3.3, released 19 January 2008)
%  Article to demonstrate format for SPIE Proceedings
%  Special instructions are included in this file after the
%  symbol %>>>>
%  Numerous commands are commented out, but included to show how
%  to effect various options, e.g., to print page numbers, etc.
%  This LaTeX source file is composed for LaTeX2e.

%  The following commands have been added in the SPIE class 
%  file (spie.cls) and will not be understood in other classes:
%  \supit{}, \authorinfo{}, \skiplinehalf, \keywords{}
%  The bibliography style file is called spiebib.bst, 
%  which replaces the standard style unstr.bst.

\documentclass[]{spie}  %>>> use for US letter paper
%%\documentclass[a4paper]{spie}  %>>> use this instead for A4 paper
%%\documentclass[nocompress]{spie}  %>>> to avoid compression of citations
%% \addtolength{\voffset}{9mm}   %>>> moves text field down
%% \renewcommand{\baselinestretch}{1.65}   %>>> 1.65 for double spacing, 1.25 for 1.5 spacing 
%  The following command loads a graphics package to include images 
%  in the document. It may be necessary to specify a DVI driver option,
%  e.g., [dvips], but that may be inappropriate for some LaTeX 
%  installations. 
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}

\title{Texture mapping 3D planar models of indoor environments with noisy camera poses}

%>>>> The author is responsible for formatting the 
%  author list and their institutions.  Use  \skiplinehalf 
%  to separate author list from addresses and between each address.
%  The correspondence between each author and his/her address
%  can be indicated with a superscript in italics, 
%  which is easily obtained with \supit{}.

\author{Peter Cheng, Michael Anderson, Stewart He, Avideh Zakhor
\skiplinehalf
University of California, Berkeley\\
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
%>>>> uncomment following for page numbers
% \pagestyle{plain}    
%>>>> uncomment following to start page numbering at 301 
%\setcounter{page}{301}

\begin{document}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{abstract}
  Automated 3D modeling of building interiors is used in applications
  such as virtual reality and environment mapping. Texturing these
  models allows for photorealistic visualizations of the data
  collected by such modeling systems. Camera poses obtained by these
  systems and used for texturing often suffer from inaccuracies
  however, resulting in visible discontinuities when successive images
  are projected onto a surface for texturing. Existing methods to
  stitch images together are often computationally expensive and work
  independently of pose estimates and geometry information. Using data
  from a human-operated backpack system, we propose an efficient
  method to refine camera poses in 2D using both existing estimates
  and geometry information, followed by two different methods to
  composite images together, based on the geometry of surfaces being
  textured. The effectiveness of our methods is demonstrated on a
  number of different indoor environments.
\end{abstract}

% >>>> Include a list of keywords after the abstract

\keywords{Texture Mapping, Reconstruction, Image Stitching, Mosaicing}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:introduction}
Three-dimensional modeling of indoor environments has a variety of
applications in fields such as architecture and simulation. Adding
accurate textures to these models is important in adding context to
digital visualizations, and allows for image-based analysis of
environments.  In this paper, we work with a backpack-mounted system,
carried by an ambulatory human. This backpack system contains a series
of laser scanners and inertial measurement units, allowing it to solve
the simultaneous localization and mapping (SLAM) problem, which
produces backpack poses over time, and a point cloud representing its
environment. For our environment geometry, we work with models
obtained by fitting low-resolution planar surfaces to these point
clouds.  While a human-carried system provides advantages over more
common wheel-mounted systems in terms of agility and portability, it
often leads to relatively high localization error even after applying
sophisticated localization techniques. Furthermore, our model geometry
often contains some error as well, and is often adjusted to ignore
low-resolution features, such as small walls, furniture, and various
small features commonly found in indoor environments. As a result,
methods for texture mapping that rely on accurate camera poses
generally produce poor results.

Our goal of texturing planar surfaces with misaligned camera poses
straddles a number of highly-researched fields, including image
stitching, panorama generation, and image-based localization
algorithims. We will examine the results of existing approaches in
Section X, and compare them with our approach, which utilizes several
aspects of our modeling system and input data.

The remainder of the paper is organized as follows. Section X demonstrates 2 existing algorithms on our datasets and discusses unique aspects of our modeling system. Section X describes the basics of image projection, and a tile-based mapping approach and its shortcomings. Section X presents our approach towards image alignment, followed by Section X, which describes our two methods of compositing images. Section X contains results and conclusions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%% References %%%%%

\bibliography{report} %>>>> bibliography data in report.bib
\bibliographystyle{spiebib} %>>>> makes bibtex use spiebib.bst

\end{document}