### 2013-01-19 spie/paper.tex
everything except geom stuff draft

%  article.tex (Version 3.3, released 19 January 2008)
%  Article to demonstrate format for SPIE Proceedings
%  Special instructions are included in this file after the
%  symbol %>>>>
%  Numerous commands are commented out, but included to show how
%  to effect various options, e.g., to print page numbers, etc.
%  This LaTeX source file is composed for LaTeX2e.

%  The following commands have been added in the SPIE class 
%  file (spie.cls) and will not be understood in other classes:
%  \supit{}, \authorinfo{}, \skiplinehalf, \keywords{}
%  The bibliography style file is called spiebib.bst, 
%  which replaces the standard style unstr.bst.

\documentclass[]{spie}  %>>> use for US letter paper
%%\documentclass[a4paper]{spie}  %>>> use this instead for A4 paper
%%\documentclass[nocompress]{spie}  %>>> to avoid compression of citations
%% \addtolength{\voffset}{9mm}   %>>> moves text field down
%% \renewcommand{\baselinestretch}{1.65}   %>>> 1.65 for double spacing, 1.25 for 1.5 spacing 
%  The following command loads a graphics package to include images 
%  in the document. It may be necessary to specify a DVI driver option,
%  e.g., [dvips], but that may be inappropriate for some LaTeX 
%  installations. 
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}

\title{Texture mapping 3D planar models of indoor environments with noisy camera poses}

%>>>> The author is responsible for formatting the 
%  author list and their institutions.  Use  \skiplinehalf 
%  to separate author list from addresses and between each address.
%  The correspondence between each author and his/her address
%  can be indicated with a superscript in italics, 
%  which is easily obtained with \supit{}.

\author{Peter Cheng, Michael Anderson, Stewart He, Avideh Zakhor
\skiplinehalf
University of California, Berkeley\\
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
%>>>> uncomment following for page numbers
% \pagestyle{plain}    
%>>>> uncomment following to start page numbering at 301 
%\setcounter{page}{301}

\begin{document}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{abstract}
  Automated 3D modeling of building interiors is used in applications
  such as virtual reality and environment mapping. Texturing these
  models allows for photo-realistic visualizations of the data
  collected by such modeling systems. We perform these tasks using a
  backpack-mounted data acquisition system. Camera poses obtained by
  this system and used for texturing often suffer from inaccuracies
  however, resulting in visible discontinuities when successive images
  are projected onto a surface for texturing. Existing methods to
  stitch images together are often computationally expensive and work
  independently of pose estimates and geometry information. By
  intelligently selecting images for texturing, we develop
  an efficient method to adjust camera poses in 2D, followed by two
  different methods to composite images together, based on the
  geometry of surfaces being textured. The effectiveness of our
  method is demonstrated on a number of different indoor
  environments.
\end{abstract}

% >>>> Include a list of keywords after the abstract

\keywords{Texture Mapping, Reconstruction, Image Stitching, Mosaicing}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:introduction} % \label{} allows reference to this section
Three-dimensional modeling of indoor environments has a variety of
applications in fields such as architecture, conservation, and
simulation. Applying accurate textures to these models is important to
add context to digital visualizations, and allows for image-based
analysis of environments.

In this paper, we perform data acquisition with a backpack-mounted
system carried by an ambulatory human \cite{liu2010indoor}. This
backpack system contains a series of cameras, laser scanners, and
inertial measurement units, used to solve the simultaneous
localization and mapping (SLAM) problem, which produces backpack (and
therefore camera) poses over time as well as a point cloud
representing the surrounding environment \cite{chen2010indoor,
  liu2010indoor, kua2012loopclosure}. For environment geometry, we
work with models obtained by fitting low-resolution planar surfaces to
these point clouds \cite{sanchez2012point}.

While a human-carried system provides advantages over more common
wheeled systems in terms of agility and portability, it often leads to
relatively high localization error even after applying sophisticated
localization techniques \cite{liu2010indoor}. Furthermore, our model
geometry usually contains some error as well, and is often tuned to
ignore low-resolution features, such as small walls, furniture, and
other minor non-planar features commonly found in indoor
environments. As a result, methods for texture mapping that rely on
accurate camera poses and projection surfaces generally produce poor
results, while more sophisticated methods to accurately align images
in 3D lead to high runtime on our typically very large datasets. In
contrast, we present a method, which given a high number of images,
searches for a subset of them that can be successfully aligned with 2D
adjustments alone.

The remainder of the paper is organized as follows. Section
\ref{sec:simpleTextureMapping} explains how images are projected onto
our geometry and explains a simple approach towards texturing. Section
\ref{sec:existingApproaches} covers more sophisticated existing
approaches, and their performance on our datasets. Section
\ref{sec:2dAlignment} contains our efficient approach towards image
alignment, followed by Section \ref{sec:imageCompositing}, which
describes our two methods of compositing images. Section
\ref{sec:resultsAndConclusions} contains results and conclusions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Simple Texture Mapping}
\label{sec:simpleTextureMapping}

The geometry of the texture mapping process for a planar surface is
shown in Figure \ref{fig:projection}.  We are provided with a set of
$M$ images to texture a target plane. Each image has a camera matrix
$P_i$ for $i=1..M$, which translates a 3D point in the world
coordinate system to a 2D point or pixel in image $i$'s coordinates. A
camera matrix $P_i$ is composed of the camera's intrinsic parameters,
containing focal length and image center, as well as extrinsic
parameters which specify the rotation and translation of the camera's
position in 3D world coordinates at the time that image $i$ is
taken. These extrinsic parameters are determined by the backpack
hardware and localization algorithms \cite{chen2010indoor,
  liu2010indoor, kua2012loopclosure} and are substantially noisy.

Because our backpack system takes photos at a rate of 5 Hz, thousands
of images are available for texturing each surface in our model. Our
goal in designing a texture mapping process is to decide which of
these images should be used, and where their contents should map onto
the final texture, in order to minimize any visual discontinuities or
seams that would suggest that the plane's texture is not composed of a
single continuous image.

\begin{figure}
  \begin{minipage}[b]{0.45\linewidth}
    \centering
    \includegraphics[height=1.5in]{Projection.pdf}
    \caption{Surfaces to be textured are specified in 3D space by
      corners $C_i$. Images are related to each plane through the
      camera matrics $P_{1..m}$. }
    \label{fig:projection}
  \end{minipage}
  \hspace{0.5cm}
  \begin{minipage}[b]{0.45\linewidth}
    \centering
    \includegraphics[height=1.125in]{scoringFunction.jpg}
    \caption{We minimize camera angle $\alpha$ and distance $d$ by
      maximizing the scoring function $\frac{1}{d} (-1 \cdot \vec{c})
      \cdot \vec{n}$}
    \label{fig:scoringFunction}
  \end{minipage}
\end{figure}

\subsection{Tile-based Texture Mapping}
\label{sec:tileBasedMapping}
Ignoring the fact that the camera matrices $P_{1..M}$ are inaccurate,
a simple texturing approach can be performed by discretizing the
target plane into small square tiles, in our case 5 pixels across, and
choosing an image to texture each tile directly.

We choose to work with rectangular units to ensure that borders
between any two distinct images in the final texture are either
horizontal or vertical. Since most strong environmental features
inside buildings are horizontal or vertical, any visible seams in our
texture will intersect them minimally and be less noticeable.

In order to select an image for texturing a tile $t$, we first gather
a list of candidate images that contain all four of its corners, which
we can quickly check by projecting $t$ into each image using the $P_i$
camera matrices. Furthermore, each candidate image must have been
taken at a time when its camera had a clear line-of-sight to $t$,
which can be calculated using standard ray-polygon intersection tests
between the camera location, $t$, and every other plane
\cite{rayintersection}.

Once we have a list of candidate images for $t$, we define a scoring
function in order to objectively select the best image for texturing
$t$. Since resolution decreases and camera pose errors become more
extreme over distance, we wish to minimize the distance between
cameras and the surfaces they texture. Additionally, we desire images
that are projected perpendicularly onto the plane, maximizing the
resolution and amount of useful texture available in their
projections, as well as minimizing any parallax effects due to
real-world geometry not accurately represented by our digital
model. In other words, we wish to minimize the angle between the
tile's normal vector and the camera axis for images selected for
texturing that tile. These two criteria can be met by maximizing the
function $\frac{1}{d} (-1 \cdot \vec{c}) \cdot \vec{n}$ as shown in
Figure \ref{fig:scoringFunction}. Specifically, $d$ is the distance
between the centers of a camera and a tile, and $\vec{n}$ and
$\vec{c}$ are the directions of the plane's normal and the camera axis
respectively.

\begin{figure}
  \centering \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_naive.jpg}}
  \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_naive_shift.jpg}}

  \centering \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_cache_shift.jpg}}
  \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_shortest_shift.pdf}}

  \centering \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_cache_shift_details.png}}
  \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_shortest_shift_details.png}}

  \centering \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_cache_shift_blend.pdf}}
  \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_shortest_shift_blend.pdf}}

  \caption{(a): Tile-based texturing. (b): Tile-based texturing after
    image alignment. (c): Tile-based texturing after image alignment
    with caching. (d): Shortest path texturing after image
    alignment). (e,f): Comparison of image artifacts in (c)
    vs. (d). (g,h): Blending applied to (c) and (d).}
  \label{fig:compareAll}
\end{figure}

As Figure \ref{fig:compareAll}(a) demonstrates, this approach seleccts
images with high resolution, and handles parallax effects well, as
seen with the table and the poster at the lower-right of the
image. Obviously, there are many image boundaries with abrupt
discontinuities between tiles, due to significant misalignment between
images, though most of it appears to be reconcilable using 2d
transforms. Given our high number of input images, a better image
selection procedure will further improve results, as described in
Section \ref{sec:imageCompositing}, but a more significant and
reliable improvement can be found through first improving the
alignment of our images.

\section{Existing Approaches to Image Alignment}
Stitching together multiple images to produce a larger, seamless image
is a commonly performed task, with many successful approaches over the
past years. Generally, parts of images are matched to each other,
usually through direct comparisons or feature detection and
matching. Images are then transformed to maximize matches, often by
calculating homographies between pairs of images, or by iteratively
adjusting camera poses in 1 to 6 degrees of freedom.

Feature matching has a number of advantages over direct matching that
make it more suitable for our often non-planar data, and rotational
differences, as covered in X. Feature matching however, works best
when multiple unique visual references exist in the environment that
can be detected in multiple images. In contrast, our indoor
environments have a high prevalence of bare surfaces, as well as
repeating textures, such as similar windows, doors, and wall-mounted
decorations, that cause difficulty in disambiguating features. This
lack of reliable reference points often results in errors when
matching images together.

Additionally, our datasets often contain long chains of images, which
leads to error accumulation when image correspondences are not exact
and are compounded together. For example, when matching a long chain
of images through homographies, a pixel in the $nth$ image is
translated into the first image's coordinates by multiplying by the
$3\times3$ matrices $H_1 H_2 H_3 ... H_n$. Errors in any of these
homography matrices are propagated to all further images, resulting in
drift.

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=6in]{autostitchResult.jpg}}

  \centering
  \subfloat[][]{\includegraphics[width=6in]{graphApproach.jpg}}

  \centering \subfloat[][]{\includegraphics[width=6in]{finalLong.jpg}}

  \caption{Texture alignment via (a) image mosaicing, (b) the
    graph-based localization refinement algorithm from
    \cite{chen2010indoor}, and (c) the method proposed in this paper.}
  \label{fig:mosaic3D}
\end{figure}

Past work with our data acquisition backpack integrated image
stitching with an iterative localization algorithm, and performed
homography-based camera pose refinement. When run on long chains of
images, especially where features are sparse, it produces distorted
textures, as seen in Figure X, due to the problems described
above. This approach is also WHATEVER NONLINEAR, and its iterative
camera adjustment process over our large datasets is not guaranteed to
SOMETHING, and leads to prohibitively long computation time.

The AutoStitch software package performs homography-based alignment as
well, with additional provisions that attempt to reduce drift (cite
X,X). Though AutoStitch also performs well in areas with dense
features, it can not handle areas without features, and has trouble
aligning wall sections with even short segments of bare texture. The
example in Figure X was generated after manual tuning, and areas with
fewer visual features or repeating texture patterns simply failed
outright with AutoStitch.

\section{2D Image Alignment with Geometry Information}
\label{sec:2dAlignment}

In this section, we describe our method of efficient and robust image
alignment. Though state-of-the-art techniques towards image stitching
work in 3D, we sacrifice THAT WORD for efficiency, and elect to
perform image alignments entirely in 2D. Besides the reduced
complexity, we have found 2D alignment to work well for two main
reasons. First, because we have such a large number of images to
choose from, we have found that the compositing approache in Section X
is able to select a subset of images that align well, given only 2D
alignments. Furthermore, the nature of our input data is such that
localization error chiefly occurs in two dimensions, in the same plane
as the surface being projected onto. Recall that our data acquisition
system is backpack-mounted, with cameras facing to the sides of the
operator. The operator makes efforts to walk parallel to walls, and
the localization and model-generation algorithms that provide our
input conform the operator's path to be straight and parallel to
detected surfaces \cite{kua2012loopclosure, sanchez2012point}. Since
the operator walks in as upright a position as possible, errors in
roll, yaw, and the operator's distance from parallel surfaces are
minimal. Thus, our highest errors stem from uncertainty in the
operator's pitch, equivalent to rotation around the camera axes, as
well as location along surfaces. These equate to 2D rotation and
translation in a surface's plane.

Our alignment procedure consists of three main parts. First, all
images are projected onto the surface and lines within these
projections are detected. These lines are then matched up with
geometry-based lines comprising the surface's boundary and
intersection with other surfaces. Second, occlusion checks are
performed to remove invalid parts of each image for the target
surface. These two steps are image-independent and performed in
parallel. Third, we detect SIFT feature matches between pairs of
images and solve a linear least squares problem in 2D to maximize
matches.

\subsection{Geometry-based Alignment}
\label{sec:geometryAlignment}
After computing each image's projection onto the target surface, as
described in Section \ref{sec:simpleTextureMapping}, we detect line
segments in the image projections using Hough transforms. Experience
and intuition show that walls in indoor environments often contain
linear features that are either horizontal or vertical, corresponding
to doors, windows, posters, etc. Thus, when texturing walls, we rotate
images in 2D such that dominant lines are made to be horizontal or
vertical. This can be further extrapolated by orienting lines with a
wall's boundaries, for example in areas with a slanted roof.

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[height=1.5in]{geometryAlign_planes.png}}
  \subfloat[][]{\includegraphics[height=1.5in]{geomAlign}}
  \caption{(a) When texturing the red surface, it makes sense to align
    images to surface boundaries and intersections with other
    surfaces. (b) The darker lines are geometry-based lines, while
    lighter lines are lines detected in the image via Hough
    transform. Above is the original projection using the input camera
    poses, and below is the projection after rotation and translation
    for alignment.}
  \label{fig:geometryAlignment}
\end{figure}

At this point in time, image occlusions have not been accounted
for. As a result, some image projections contain texture that should
project to an adjacent surface, generally with a linear boundary where
the two surfaces meet. If this linear boundary is detected by Hough
transform, the image is rotated and shifted in 2D such that the visual
boundary between two surfaces in an image projection matches the
physical boundary in our digital model. An example of such an
adjustment is in Figure \ref{fig:geometryAlignment}

To perform the rotation and alignment, we use the RANSAC
\cite{fischler1981random} framework. RANSAC allows us to determine an
optimal rotation and translation to match up two groups of line
segments, while accounting for outliers. First, a rotation angle is
computed by gathering all pairs of image line segments and geometry
line segments with less than 20 degrees between them. We then run the
RANSAC algorithm by sampling 4 of these angles at a time, and using
their average as the fitting function. Angles over 2 degrees from this
average are considered outliers. Once this 2d rotation angle is
determined, it is applied to the image, and projections are
re-calculated and image-based lines are re-detected. Now, all pairs of
image line segments and geometry line segments with less than 0.2
degree difference and less than 250 mm distance at their furthest
points are gathered. RANSAC is then similarly employed to calculate
optimal horizontal and vertical translations independently, by
sampling 4 pairs at a time and considering an outlier threshold of 10
mm. If a translation is applied, we indicate that the image was
aligned to geometry in one or both dimensions, which will reduce
further shifting in Section \ref{sec:robustSIFTFeatureMatching}.

\subsection{Image Occlusion}
\label{sec:imageOcclusion}
Now that images have been aligned to geometry where possible, we
perform a simple recursive occlusion procedure on each image. This is
done by performing the intersection tests in section
\ref{sec:simpleTextureMapping} in a regularly spaced grid. Where four
corners of a rectangular region are occluded, texture is
removed. Where no corners are occluded, nothing occurs. Where there is
a mixture of both, the rectangular region is subdivided into four, and
the same process is performed on each. By performing Section
\ref{sec:geometryAlignment}'s alignment procedure before occlusion,
texture belonging to other surfaces is accurately removed, which is
necessary for the next section.

\subsection{2D Feature Alignment}
\label{sec:robustSIFTFeatureMatching}
Our next step is to align overlapping images by searching for
corresponding points between all pairs of overlapping images. We use
SIFT features for their high detection rate, and choose to use feature
alignment rather than pixel or intensity-based alignment due to the
differences in lighting as well as possible occlusion among our
images, both of which feature alignment is less sensitive to
\cite{lowe1999object, mikolajczyk2005performance, szeliski2006image}.
We use SiftGPU \cite{siftgpu} for its high performance on both feature
detection as well as pairwise matching. These matches determine $d^x$
and $d^y$ distances between each pair of features for two image
projections, though these distances may not always be the same for
different features. Since indoor environments often contain repetitive
features such as floor tiles or doors, we need to ensure that
SIFT-based distances are reliable. First, we only perform alignment on
parts of images that overlap given the original noisy poses. Second,
we discard feature matches that correspond to an image distance
greater than 40 pixels from what the noisy poses estimate. In order to
utilize the remaining feature matches robustly, RANSAC
\cite{fischler1981random} is again used to estimate the optimal
$d^x_{i,j}$ and $d^y_{i,j}$ distances between two images $i$ and
$j$. For this application, the RANSAC fitting function finds the
average distance between selected matches in a pair of images, and the
distance function for a pair of points is chosen to be the difference
between those points' SIFT match distance and the average distance
computed by the fitting function. We use a 10 pixel threshold, so that
SIFT matches are labeled as outliers if their horizontal or vertical
distances are not within 10 pixels of the average distance computed by
the fitting function.

We now use the RANSAC-calculated $d^x_{i,j}$ and $d^y_{i,j}$ distances
between each pair of images to refine their positions using weighted
linear least squares. There are a total of $M^{2}$ possible pairs of
images, though we only generate distances between images with SIFT
matches. Given these distances and the original image location
estimates, we can solve a least squares problem
($\textrm{min}_{\vec{\beta}} ||A \vec{\beta} - \vec{\gamma}||_2^2 $)
to estimate the location of the images on the plane. The
$M$-dimensional vector $\vec{\beta}$ represents the unknown $x$
location of each image on the plane for $1 \dots M$. The optimal $x$
and $y$ locations are obtained in the same way, so we only consider
the $x$ locations here:

\[\vec{\beta} =
\begin{pmatrix}
  x_1, & x_2, & x_3, & \cdots & x_{M-1}, & x_M
\end{pmatrix}
\]

The $N \times M$ matrix $A$ is constructed with one row for each pair
of images with measured distances produced by the SIFT matching
stage. A row in the matrix has a $-1$ and $1$ in the columns
corresponding to the two images in the pair. For example, the matrix
below indicates a SIFT-based distance between images 1 and 2, images 1
and 3, images 2 and 3, etc.
\[
A =
\begin{pmatrix}
  -1 & 1 & 0 & \cdots & 0 & 0\\
  -1 & 0 & 1 & \cdots & 0 & 0\\
  0 & -1 & 1 & \cdots & 0 & 0\\
  \vdots  & \vdots & \vdots & \ddots & \vdots  & \vdots\\
  0 & 0 & 0 & \cdots & 1 & 0 \\
  0 & 0 & 0 & \cdots & -1 & 1 \\
  1 & 0 & 0 & \cdots & 0 & 0\\
\end{pmatrix}
\]
If only relative distances between images are included, the absolute
location of the images can not be calculated, and the matrix is rank
deficient. From Section \ref{sec:geometryAlignment}, a number of
images are anchored to geometry points, in one or both dimensions, and
thus their locations can be used to fix the rest in place. In case no
anchor images exist, we simply arbitrarily pick an image, as we have
no other reference points to work with. In the above matrix, the first
image is set to be such an anchor, simply by placing a $1$ in its
column.

The $N$-dimensional observation vector $\vec{\gamma}$ is constructed
using the SIFT-based distances generated in the RANSAC matching
stage. Elements in the observation vector corresponding to anchor
images are simply their locations as determined by the original noisy
localization. Thus a $\vec{\gamma}$ corresponding to the above matrix
can be written as:

\[
\vec{\gamma}^T =
\begin{pmatrix}
  d_{1,2}, &d_{1,3}, &d_{2,3}, &\hdots &d_{N-2,N-1}, &d_{N-1,N}, &x_1
\end{pmatrix}
\]

The $\vec{\beta}$ that minimizes $||A \vec{\beta} -
\vec{\gamma}||_2^2$ results in a set of image locations on the plane
that best honors all the SIFT-based distance measurements between
images. This solution however does not make use of our noisy camera
poses, and will fail when no SIFT matches are found between one
segment of the plane and another. To account for this, we add rows to
the $A$ matrix and observations to the $\vec{\gamma}$ vector
corresponding to the original noisy distances. We then add weighting
to our least squares problem where the SIFT distances and anchor
values are given a high weight e.g. 1, while the noisy distances are
given a smaller weight e.g. 0.01.

After completing this same process for the $y$ dimension as well, and
making the resultant shifts, our images overlap and match each other
with far greater accuracy. Applying the simple mapping scheme in
Section \ref{sec:tileBasedMapping} to the same wall used in that
section results in Figure \ref{fig:compareAll}(b), which has far fewer
discontinuities, though errors due to lighting differences and
repeating features are still visible.

\section{Image Compositing}
\label{sec:imageCompositing}
In this section, we revisit the tile-based texturing approach from
Section \ref{sec:simpleTextureMapping}, with an added caching
mechanism to reduce image boundaries. This method works well given all
manner of camera poses and surfaces, but for optimal cases where we
have large sections of usable texture from images, we propose a
superior method that further reduces image boundaries.

\subsection{Tile-Mapping with Caching}
\label{sec:mappingWithCaching}
From Section \ref{sec:tileBasedMapping}, we saw that discontinuities
occur where adjacent tiles are textured by different images. Though
Section \ref{sec:2dAlignment}'s image alignment removes many such
discontinuities, there are still cases where seams are visible due to
imprecise matching or other factors such as model-based errors. To
reduce the cases where this happens, it makes sense to take into
account image selections made by neighboring tiles while texture
mapping a given tile. By using the same image across tile boundaries,
we can eliminate a discontinuity altogether. If this is not possible
because a tile is not visible in images used by neighboring tiles,
using similar images across tile boundaries also leads to less
noticeable discontinuities.

Essentially a caching mechanism, we select the best image for a tile
$t$ by searching through two subsets of images for a viable candidate,
before searching through the entire set of valid images. The first
subset of images is the images selected by adjacent tiles that have
already been textured. We must first check which of these images can
map to $t$, and then of those, we make a choice according to the
scoring function in Figure \ref{fig:scoringFunction}. Before reusing
this image, we ensure it meets the criteria $\alpha < 45^\circ$, in
order to ensure a high resolution projection, with $\alpha$ as the
camera angle as shown in Figure \ref{fig:scoringFunction}.

If no satisfactory image is found in the first subset, we check a
second subset of images, consisting of those taken near the ones in
the first subset, both spatially and temporally. These images are not
the same as the ones used for neighboring tiles, but are taken at a
similar location and time, suggesting that their localization and
projection are quite similar, and thus likely matched more cleanly. If
no viable image is found according to the same criteria as before, we
search the entire set of candidate images, selecting based on the same
scoring function from Figure \ref{fig:scoringFunction}.

The result of this caching approach is shown in Figure
\ref{fig:compareAll}(c), where seams are now reduced as compared to
Figure \ref{fig:compareAll}(b). However, some discontinuities are
still present, as visible in the many posters on the wall with breaks
in their borders.

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=1in]{projectionWall.pdf}}
  \centering
  \subfloat[][]{\includegraphics[width=1.1in]{projectionCeiling.pdf}}
  \centering
  \subfloat[][]{\includegraphics[width=0.9in]{projectionWallCrop.pdf}}
  \caption{(a) Images for vertical planes are tilted, but their camera
    axes are more or less normal to their respective planes. (b)
    Camera axes for ceiling images are at large angles with respect to
    plane normals. (c) Wall images are cropped to be rectangular.}
  \label{fig:projectionAngles}
\end{figure}

As mentioned earlier, our data comes from a mobile backpack
system. Human operators can not carry the backpack in a perfectly
upright position and are bent forwards at 15 to 20 degrees with
respect to the vertical direction. As a result, cameras facing
sideways are head on with respect to vertical walls, while cameras
oriented towards the top or bottom of the backpack are at an angle
with respect to horizontal floors and ceilings. This is depicted in
Figures \ref{fig:projectionAngles}(a) and
\ref{fig:projectionAngles}(b). These oblique camera angles for
horizontal surfaces translate into textures that span large areas once
projected, as shown in Figure \ref{fig:projectionAngles}(b). Using the
tile-based texture mapping criteria from Figure
\ref{fig:scoringFunction}, such projections have highly varying scores
depending on the location of a tile on the plane. Thus, the tiling
approach in this section is a good choice for texturing floors and
ceilings, as it uses the parts of image projections that maximize
resolution and accuracy for their respective plane locations,
e.g. areas near point A and not near point B, in Figure
\ref{fig:projectionAngles}(b).

\subsection{Shortest Path Texturing}
\label{sec:shortestPath}
For vertical walls, most images are taken from close distances and
head-on angles, resulting in high resolution fronto-parallel
projections. As a result, for each tile on a wall plane, the scoring
function of Figure \ref{fig:scoringFunction} is relatively flat with
respect to candidate images, as they are all more or less head
on. Thus, the scoring function is less significant for walls, and it
is conceivable to use a different texturing strategy to directly
minimize visible seams when texturing them. This is done by choosing
the smallest possible set of images that (a) covers the entire plane
and (b) minimizes the visibility of borders between them. A
straightforward cost function that accomplishes the latter is the sum
of squared differences (SSD) of pixels in overlapping regions between
all pairs of images. Minimizing this cost function encourages image
boundaries to occur either in featureless areas, such as bare walls,
or in areas where images match extremely well.

The first step in this procedure is to obtain a list of images useful
for texturing the wall. A simple way to do this is to use the images
selected by the tiling process in Section
\ref{sec:tileBasedMapping}. Such a list of images is guaranteed to
cover the entire wall, and consists of desired camera poses overall.

To cover the entirety of a plane, our problem can be defined as
minimally covering a polygon i.e. the planar surface, using other
polygons of arbitrary geometry i.e. image projections, with the added
constraint of minimizing the cost function between chosen images.
This is a complex problem, though we can take a number of steps to
simplify it.

Given that wall-texture candidate images are taken from more or less
head-on angles, and knowing that only minor rotations are made in
Section \ref{sec:2dAlignment}, we can reason that their projections
onto the plane are approximately rectangular. By cropping them all to
be rectangular, as shown in Figure \ref{fig:projectionAngles}(c), our
problem becomes the conceptually simpler one of filling a polygon with
rectangles, such that the sum of all costs between each pair of
rectangles is minimal. We thus also retain the advantages of working
with rectangular units, as explained in Section
\ref{sec:tileBasedMapping}.

The operator's path, and the location and orientation of the cameras
on the acquisition backpack is such that images nearly always contain
the entirety of the floor to ceiling range of wall planes. Images
therefore rarely need to be projected with one above another when
texturing wall planes. In essence, we need only to ensure lateral
coverage of wall planes, e.g. from left to right, as our images
provide full vertical coverage themselves. We can thus construct a
Directed Acyclic Graph (DAG) from the images, with edge costs defined
by the SSD cost function, and solve a simple shortest path problem to
find an optimal subset of images with regard to the SSD cost function
\cite{dijkstra}.

\begin{figure}
  \centering
  \includegraphics[width=3in]{dagCreation.pdf}
  \caption{DAG construction for the image selection process. \\}
  \label{fig:dagCreation}
\end{figure}

Figure \ref{fig:dagCreation} demonstrates the construction of a DAG
from overlapping images of a hallway wall. Images are sorted by
horizontal location left to right, and become nodes in a
graph. Directed edges are placed in the graph from left to right
between overlapping images. The weights of these edges are determined
by the SSD cost function. Next, we add two artificial nodes, one start
node representing the left border of the plane, and one end node
representing the right border of the plane. The left(right) artificial
node has directed edges with equal and arbitrary cost $C_0$ to(from)
all images that meet the left(right) border of the plane. We now solve
the shortest path problem from the start node to the end node. This
results in a set of images completely covering the plane horizontally,
while minimizing the cost of seams between images.

In rare cases where the vertical dimension of the plane is not
entirely covered by one or more chosen images, we are left with holes
where no images are selected to texture. Since these holes are rare,
and generally fairly small, we use a greedy approach, repeatedly
filling the hole with images that result in the lowest SSD costs in a
blending region around the hole, as discussed in Section
\ref{sec:blending}. This method is not as optimal as a true
2D-coverage solution would be, but it is a fast approximation, and
adequately handles the few holes we encounter.

With this completed, we have now mapped every location on the plane to
at least one image, and have minimized the number of images, as well
as the discontinuities at their borders. As seen in Figure
\ref{fig:compareAll}(d), this shortest path method has fewer visible
discontinuities than Figure \ref{fig:compareAll}(c) corresponding to
the tile caching approach\footnote{In Figure \ref{fig:compareAll}(d),
  we arbitrarily chose one image for texturing where images overlap,
  as blending will be discussed in section \ref{sec:blending}.}. This
is especially evident when comparing the posters in the images, as
seen in Figures \ref{fig:compareAll}(e) and
\ref{fig:compareAll}(f). This shortest path approach approach directly
reduces the cost of each image boundary, while the tile caching method
uses a scoring function that only approximates this
effect. Furthermore, this approach guarantees the best selection of
images to minimize seams, while the sequential tile caching method may
select images early on that turn out to be poor choices once
subsequent tiles have been processed. This shortest path approach is
also far less intensive in terms of memory usage and runtime, both
during texture generation and rendering, as it does not require
discretizing planes or images.

When texturing an entire 3D planar model, we apply the shortest path
method on walls, due to its superior output when provided with head-on
images. Floors and ceilings however, given their many images taken at
oblique angles, are textured using the tile caching method.

\subsection{Blending}
\label{sec:blending}
We now apply a blending procedure to both texturing methods. Although
the image alignment steps and image selection in both methods attempt
to minimize all mismatches between images, there are occasional
unavoidable discontinuities in the final texture due to different
lighting conditions or inaccuracies in model geometry. These can
however be treated and smoothed over by applying alpha blending over
image seams.  Whether the units we are blending are
rectangularly-cropped images or rectangular tiles, we can apply the
same blending procedure, as long as we have a guaranteed overlap
between units to blend over.

For the tile caching method, we can ensure overlap by texturing a
larger tile than needed for display. For example, for a rendered tile
$l_1 \times l_1$, we can associate it with a texture $(l_1 + l_2)
\times (l_1 + l_2)$ in size.  We have found $l_2 = \frac{l_1}{2}$ to
provide a balance between blending and keeping features unblurred. For
the shortest path method, we have already ensured overlap between
images. To enforce consistent blending however, we add a minimum
required overlap distance of 40 px while solving the shortest path
problem in Section \ref{sec:shortestPath}. Additionally, if images
overlap in a region greater than the overlap distance, we only apply
blending over an area equal to the overlap distance.

After performing linear alpha blending across overlapping regions, our
texture mapping process is complete. Figures \ref{fig:compareAll}(g)
and \ref{fig:compareAll}(h) show the blended versions of Figures
\ref{fig:compareAll}(c) and \ref{fig:compareAll}(d) respectively. It
is clear that the shortest path approach exhibits much better
alignment and fewer seams than the tile-caching method, as Figure
\ref{fig:compareAll}(h) clearly has the best visual quality among the
textures in Figure \ref{fig:compareAll}.

\section{Results and Conclusions}
\label{sec:resultsAndConclusions}
Examples of ceilings and floors textured with the tile caching
approach, and walls textured with the shortest path approach, are
displayed in Figure \ref{fig:results}. High resolution colored texture
comparisons, as well as further examples and interactive walkthroughs
are available at
\footnote{\url{http://www.eecs.berkeley.edu/~pcheng/indoormapping}}.

As mentioned earlier, our approach is quite efficient. The top wall in
Figure \ref{fig:results}(a) was generated with 7543 $\times$ 776
pixels, and spans a 40-meter long wall. Given 41000 input images, a
2.8GHz consumer-grade laptop takes approximately a minute to pick 36
candidate images, followed by another minute to perform both image
alignment and the shortest path texturing method, though over 75\% of
that time is spent calculating SIFT matches within the SiftGPU
framework. While not real-time, the process is capable of generating
quick updates after changes in various parameters or modifications to
input data, and if integrated directly into a modeling system, could
provide live visualization as data is collected.

In this paper, we have developed an approach to texture mapping models
with noisy camera localization data. We are able to refine image
locations based on geometry references and feature matching, and
robustly handle outliers. Using the tile-based mapping approach, we
can texture both simple rectangular walls as well as complex floor and
ceiling geometry. We also implemented a shortest path texturing method
that produces seamless textures on planes where multiple head-on
images are available. Both of these approaches are highly modular, and
easily tunable for different environments and acquisition systems.

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=3in]{finalfloors.jpg}} ~~~~~~~~
  \centering
  \subfloat[][]{\includegraphics[width=3in]{finalceilings.jpg}}

  \centering \subfloat[][]{
    \includegraphics[height=2in, width=3in]{floorcropped.jpg} ~~~~~~~~
    \includegraphics[width=3in]{pier15floor.jpg}
  }

  \centering{
    \includegraphics[width=1.9in]{fullmodel.png}
    \includegraphics[width=1.9in]{threestoryfull2.png}
    \includegraphics[width=1.9in]{pier15.png} \\
    \includegraphics[width=6in]{threestoryfull.png}
  }

  \centering \subfloat[][]{ }
  \caption{Examples of our final texture mapping output for (a) walls,
    (b) ceilings, (c) floors, (d) full models.}
  \label{fig:results}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%% References %%%%%

\bibliography{report} %>>>> bibliography data in report.bib
\bibliographystyle{spiebib} %>>>> makes bibtex use spiebib.bst

\end{document}