### 2013-02-04 spie/paper.tex
done with text-based edits

%  article.tex (Version 3.3, released 19 January 2008)
%  Article to demonstrate format for SPIE Proceedings
%  Special instructions are included in this file after the
%  symbol %>>>>
%  Numerous commands are commented out, but included to show how
%  to effect various options, e.g., to print page numbers, etc.
%  This LaTeX source file is composed for LaTeX2e.

%  The following commands have been added in the SPIE class 
%  file (spie.cls) and will not be understood in other classes:
%  \supit{}, \authorinfo{}, \skiplinehalf, \keywords{}
%  The bibliography style file is called spiebib.bst, 
%  which replaces the standard style unstr.bst.

\documentclass[]{spie}  %>>> use for US letter paper
%%\documentclass[a4paper]{spie}  %>>> use this instead for A4 paper
%%\documentclass[nocompress]{spie}  %>>> to avoid compression of citations
%% \addtolength{\voffset}{9mm}   %>>> moves text field down
%% \renewcommand{\baselinestretch}{1.65}   %>>> 1.65 for double spacing, 1.25 for 1.5 spacing 
%  The following command loads a graphics package to include images 
%  in the document. It may be necessary to specify a DVI driver option,
%  e.g., [dvips], but that may be inappropriate for some LaTeX 
%  installations. 
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{float}
\title{Texture mapping 3D planar models of indoor environments with noisy camera poses}

%>>>> The author is responsible for formatting the 
%  author list and their institutions.  Use  \skiplinehalf 
%  to separate author list from addresses and between each address.
%  The correspondence between each author and his/her address
%  can be indicated with a superscript in italics, 
%  which is easily obtained with \supit{}.

\author{Peter Cheng, Michael Anderson, Stewart He, Avideh Zakhor
\skiplinehalf
University of California, Berkeley\\
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
%>>>> uncomment following for page numbers
% \pagestyle{plain}    
%>>>> uncomment following to start page numbering at 301 
%\setcounter{page}{301}

\begin{document}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{abstract}
  Automated 3D modeling of building interiors is used in applications
  such as virtual reality and environment mapping. Texturing these
  models allows for photo-realistic visualizations of the data
  collected by such modeling systems. While data acquisition times for
  mobile mapping systems are considerably shorter than for static
  ones, their recovered camera poses often suffer from inaccuracies,
  resulting in visible discontinuities when successive images are
  projected onto a surface for texturing. Existing methods to stitch
  images together are often computationally expensive and work
  independently of pose estimates and geometry data. We present a
  method for texture mapping that reduces complexity by selecting
  images whose camera poses can be align in 2D. We perform alignment
  of images to geometry as well as images to each other, which
  produces visually consistent textures even in the presence of
  inaccurate surface geometry. We also present two different methods
  for selecting and compositing images into a single texture,
  depending on the geometry of the surface being textured. The
  effectiveness of these methods is demonstrated on a number of different
  indoor environments.
\end{abstract}

% >>>> Include a list of keywords after the abstract

\keywords{Texture Mapping, Reconstruction, Image Stitching, Mosaicing}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:introduction} % \label{} allows reference to this section
Three-dimensional modeling of indoor environments has a variety of
applications such as training and simulation for disaster management,
virtual heritage conservation, and mapping of hazardous sites. Manual
construction of these digital models can be time consuming, and as
such, automated 3D site modeling has garnered much interest in recent
years.

The first step in automated 3D modeling is the physical scanning of
the environment's geometry. An indoor modeling system must be able to
recover camera poses within an environment while simultaneously
reconstructing the 3D structure of the environment itself
\cite{chen2010indoor, hz, kua2012loopclosure, liu2010indoor}. This is
known as the simultaneous localization and mapping (SLAM) problem, and
is generally solved by taking readings from laser range scanners,
cameras, and inertial measurement units (IMUs) at multiple locations
within the environment.

Mounting such devices on a platform carried by an ambulatory human
provides unique advantages over vehicular-based systems on wheels in
terms of agility and portability, but can also result in larger
localization error \cite{liu2010indoor}. As a result, common methods
for texture mapping generally produce poor results.

In this paper, we present a number of approaches to texture mapping 3D
models of indoor environments made of planar surfaces in the presence
of uncertainty and noise in camera poses. In particular, we consider
data obtained from a human-operated backpack system with a number of
laser range scanners as well as 2 cameras facing left and right, each
equipped with fisheye lenses reaching an approximately 180$^{\circ}$
field of view and taking photos at a rate of 5 Hz. Applying multiple
localization and loop-closure algorithms on the raw data collected by
the onboard sensors \cite{chen2010indoor, kua2012loopclosure,
  liu2010indoor}, the backpack is localized\footnote{In this paper, we
  use the terms localization and pose recovery interchangeably, in
  that they both refer to recovering position and orientation.}  over
its data collection period. This requires recovering the 6 degrees of
freedom for the backpack as well as the cameras rigidly mounted on
it. Once this is complete, the data from the laser range scanners is
used to generate a 3D point cloud of the surrounding environment, from
which a 3D planar model is created \cite{sanchez2012point}. This
model, consisting of 2D polygonal planes in 3D space, along with the
set of images captured by the backpack's cameras and their noisy 3D
poses, can be considered the input to our texture mapping problem.

We perform texture mapping on each planar surface independently and in
parallel. For each surface, we begin by obtaining a set of images that
spans the entire surface with high resolution imagery. We then use our
noisy camera poses to project these selected images onto the
surface. These projections are then refined in 2D, in order to
maximally align them with the surface's geometry, as well as to each
other, allowing us to handle both errors in geometry as well as camera
poses. For surfaces and camera poses at any location or orientation,
we demonstrate a simple tile-based approach for sampling
high-resolution portions of images and compositing them into a
texture. In cases where cameras have consistently perpendicular
viewing angles, we demonstrate a superior approach that leads to more
seamless textures.

The remainder of the paper is organized as follows. Section
\ref{sec:simpleTextureMapping} explains how images are projected onto
geometry and shows a simple texture mapping procedure. Section
\ref{sec:existingApproaches} covers existing approaches to image
stitching, and their performance on our datasets. Section
\ref{sec:2dAlignment} contains our approach towards efficient 2D image
alignment, followed by Section \ref{sec:imageCompositing}, which
describes two methods of selecting and compositing images to create
the final texture. Section \ref{sec:resultsAndConclusions} contains
results and conclusions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Simple Texture Mapping}
\label{sec:simpleTextureMapping}

Before describing our texture mapping procedure, we use this section
to explain basic texture mapping and a method for selecting images for
texturing. The geometry of the texture mapping process for a planar
surface is shown in Figure \ref{fig:projection}.  Given a set of $M$
images to texture a target plane, camera matrix $P_i$ for the $ith$
image, translates a 3D point in the world coordinate system to a 2D
point or pixel in image $i$'s coordinates. A camera matrix $P_i$ is
composed of the camera's intrinsic parameters, containing focal length
and image center, as well as extrinsic parameters which specify the
rotation and translation of the camera's position in 3D world
coordinates at the time that image $i$ is taken. These extrinsic
parameters are determined by the backpack hardware and localization
algorithms \cite{chen2010indoor, liu2010indoor, kua2012loopclosure}
and are substantially noisy.

Because the backpack system takes pictures at a rate of 5 Hz, hundreds
of images are available for texturing each surface in the model. Our
objective in designing a texture mapping process is to determine which
of these images should be used, and where their contents should map
onto the final texture, in order to minimize any visual
discontinuities or seams that would suggest that the plane's texture
is not composed of a single continuous image. In the remainder of this
section, we provide an overview of a simple tile-based texture mapping
procedure.

\begin{figure}
  \begin{minipage}[b]{0.45\linewidth}
    \centering
    \includegraphics[height=1.8in]{Projection.pdf}
    \caption{Surfaces to be textured are specified in 3D space by
      corners $C_i$. Images are related to each surface through the
      camera matrices $P_{1..m}$. }
    \label{fig:projection}
  \end{minipage}
  \hspace{0.5cm}
  \begin{minipage}[b]{0.45\linewidth}
    \centering
    \includegraphics[height=1.8in]{scoringFunction.jpg}
    \caption{We minimize camera angle $\alpha$ and distance $d$ by
      maximizing the scoring function $\frac{1}{d} (-1 \cdot \vec{c})
      \cdot \vec{n}$}
    \label{fig:scoringFunction}
  \end{minipage}
\end{figure}

Ignoring the fact that the camera matrices $P_{1..M}$ are inaccurate,
the most straightforward approach is to discretize the target plane
into small square tiles, and choose an image to texture each tile
directly. We choose to work with rectangular units to ensure that
borders between any two distinct images in the final texture are
either horizontal or vertical. Since most strong environmental
features inside buildings are horizontal or vertical, any visible
seams in the texture intersect them minimally and are less noticeable.

In order to select an image for texturing a tile $t$, we first gather
a list of candidate images that contain all four of its corners, which
we can rapidly check by projecting $t$ into each image using the $P_i$
camera matrices. Furthermore, each candidate image must have been
taken at a time when its camera had a clear line-of-sight to $t$,
which can be determined using standard ray-polygon intersection tests
between the camera location, $t$, and every other surface
\cite{rayintersection}.

Once we have a list of candidate images for $t$, we define a scoring
function in order to objectively select the best image for texturing
$t$. Since resolution decreases and camera pose errors become more
pronounced with distance, we wish to minimize the distance between
cameras and the surfaces they texture. Additionally, we desire images
that are projected perpendicularly, rather than obliquely, onto the
plane, maximizing the resolution and amount of useful texture
available in their projections, as well as minimizing any parallax
effects due to real-world geometry not accurately represented by the
digital model. In other words, we wish to minimize the angle between
the tile's normal vector and the camera axis for images selected for
texturing that tile. These two criteria can be met by maximizing the
function $\frac{1}{d} (-1 \cdot \vec{c}) \cdot \vec{n}$ as shown in
Figure \ref{fig:scoringFunction}, where $d$ is the distance between
the centers of a camera and a tile, and $\vec{n}$ and $\vec{c}$ are
the directions of the plane's normal and the camera axis respectively.

As Figure \ref{fig:compareAll}(a) demonstrates, this approach results
in image boundaries with abrupt discontinuities between tiles, due to
significant misalignment between images, though most of it appears to
be reconcilable using 2D transforms. Given the large number of input
images, it is possible to develop a more sophisticated image selection
procedure, as described in Section \ref{sec:imageCompositing},
however, a more significant and reliable improvement can be found
through first improving the alignment of the images as shown in the
next 2 sections.

\begin{figure}[H]
  \centering \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_naive.jpg}}
  \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_naive_shift.jpg}}

  \centering \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_cache_shift.jpg}}
  \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_shortest_shift.pdf}}

  \centering \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_cache_shift_blend.pdf}}
  \subfloat[][]{\includegraphics[width=3.4in,
    height=1.2in]{wall2_shortest_shift_blend.pdf}}

  \centering \subfloat[][]{\includegraphics[width=3.4in,
    height=3.4in]{wall2_cache_comparison.png}}
  \subfloat[][]{\includegraphics[width=3.4in,
    height=3.4in]{wall2_shortest_comparison.png}}

  \caption{(a): Tile-based texturing. (b): Tile-based texturing after
    image alignment. (c): Tile-based texturing after image alignment
    with caching. (d): Shortest path texturing after image
    alignment). (e,f): Blending applied to (c) and (d). (g,h): Zoomed
    in views of discontinuities in (e) vs. in (f).}
  \label{fig:compareAll}
\end{figure}

\section{Existing Approaches to Image Alignment}
\label{sec:existingApproaches}
Stitching together multiple images to produce a larger, seamless image
is a commonly performed task, with many successful approaches over the
past years. Generally, parts of images are matched to each other,
usually through direct comparisons or feature detection and
matching. Images are then transformed to maximize matches, often by
calculating homographies between pairs of images, or by iteratively
adjusting camera poses in 1 to 6 degrees of freedom.

Feature matching has a number of advantages over direct matching that
make it more suitable for our often non-planar data, and rotational
differences \cite{szeliski2006image}. Feature matching however, works
best when multiple unique visual references exist in the environment
that can be detected in multiple images. In contrast, indoor
environments have a high prevalence of bare surfaces, as well as
repeating textures, such as similar windows, doors, and wall-mounted
decorations, that cause difficulty in disambiguating features. This
lack of reliable reference points often results in errors when
matching images together.

Additionally, our datasets often contain long chains of images,
corresponding to long hallways and corridors as shown in Figure
\ref{fig:mosaic3D}, which leads to error accumulation when image
correspondences are not accurate. For example, when matching a long
chain of images through homography, a pixel in the $nth$ image must be
translated into the first image's coordinates by multiplying by the
$3\times3$ matrix $H_1 H_2 H_3 ... H_n$. Any error in one of these
homography matrices is propagated to all further images, resulting in
drift.

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=6in]{graphApproach.jpg}}

  \centering
  \subfloat[][]{\includegraphics[width=6in]{autostitchResult.jpg}}

  \centering \subfloat[][]{\includegraphics[width=6in]{finalLong.jpg}}

  \caption{Texture alignment via (a) the graph-based localization
    refinement algorithm, (b) the AutoStitch software package, and (c)
    the proposed method.}
  \label{fig:mosaic3D}
\end{figure}

In prior work, we experimented with integrating image stitching with
an iterative localization algorithm \cite{liu2010indoor}. When run on
long chains of images, especially where features are sparse, this
approach produces distorted textures, as seen in Figure
\ref{fig:mosaic3D}(a). Furthermore, this approach is not closed-form,
and its iterative camera adjustment process over large datasets led to
prohibitively long computation time.

The AutoStitch software package performs homography-based alignment as
well, with additional provisions that attempt to reduce drift and
increase efficiency \cite{panorama2d, autostitch}. Though AutoStitch
also performs well in areas with dense features, it can not handle
areas without features, and has trouble aligning wall sections with
even short segments of bare texture. The example in Figure
\ref{fig:mosaic3D}(b) was generated after many rounds of manual
tuning, and areas with fewer visual features or repeating texture
patterns simply failed outright with AutoStitch.

\section{2D Image Alignment}
\label{sec:2dAlignment}
In this section, we describe our method for efficient and robust image
alignment. Rather than attempt to register all of our images in 3D, as
many state-of-the-art techniques for image stitching do, we instead
select a subset of images that can be aligned in 2D. The subset of
images we select for alignment and subsequent texturing corresponds to
the images selected by the tile-based texturing procedure described in
Section \ref{sec:simpleTextureMapping}.

We have found that applying 2D alignments to this set of images works
well for a few reasons. First, the nature of our input data and the
selected images is such that localization error chiefly occurs in two
dimensions, which correspond to the plane of the surface being
projected onto. This is because our backpack operator, during data
acquisition, makes efforts to walk within a few meters of, and
parallel to all walls being scanned. As a result, the translational
error of camera poses is quite minimal in the dimension perpendicular
to the surface being textured. Furthermore, because our set of images
was selected to have minimal distance from our surface, rotational
errors have a lesser effect, except for rotation around the axis
perpendicular to the surface being textured. Therefore, the majority
of both our translation and rotational errors occur in the 2
dimensions of the surface's plane. This of course does not apply for
ceilings and floors; however, the lack of features on most ceilings
and floors means that 3D misalignment does not often have a visible
effect, beyond what can be fixed in 2D.

Our proposed 2D alignment procedure consists of three parts. First, images
are projected onto the surface and lines within these projections are
detected. These lines are then used to transform images such that they match geometry-based lines composing the boundaries of the surface being textured. Second, occlusion checks are performed to remove invalid parts of each image for the target surface. Third, SIFT feature matches are detected between pairs of images, and a weighted linear least squares problem is solved in order to maximize all image and geometry-based alignments. Each step will now be explained in detail.

\subsection{Geometry-based Alignment}
\label{sec:geometryAlignment}

\begin{figure}
  \centering
  \caption{}
  \label{fig:geometryAlignment}
\end{figure}

After computing each image's projection onto the target surface, as
described in Section \ref{sec:simpleTextureMapping}, we obtain a set
of image-based line segments by using Hough transforms to detect lines
in the image projections. We then gather a set of geometry-based
lines, which correspond to the lines comprising the target surface's
border, as well as lines formed where other surfaces intersect the
target surface. An example of these lines is shown for a ceiling
surface in Figure \ref{fig:geometryAlignment}. Given perfect camera
poses and surface geometry, the lines in images corresponding to
corners between surfaces should match up exactly with borders and
intersections of surfaces. By inducing such lines to match, we fit our
camera poses more accurately to our surface, and therefore to
eachother.

To perform geometry-based alignment, we collect pairs of image-based
and geometry-based lines, which are within a distance and orientation
difference threshold of each other. We have found a distance threshold
of 250 mm and an orientation difference threshold of $10^\circ$ to
work well for our datasets. In practice, we rarely have more than 2
such pairs, but if so, we select the 2 noncollinear pairs with the
most similar orientation difference and continue. With 2 noncollinear
pairs, we select the pair with the longer image-based line, and
perform a rotation of the image such that the lines in that pair
become parallel. We then perform a translation such that the lines in
that same pair overlap. This translation has ambiguity in the
dimension along that line, which is resolved by matching the midpoint
of the image-based line in the second pair to its corresponding
geometry-based line. Thus, with 2 or more pairs, we can obtain a fixed
rotation and translation, which we save for usage in Section
\ref{sec:robustSIFTFeatureMatching}. If we only have 1 pair, we
perform a rotation and a minimal translation to match its lines. This
translation's ambiguity however, can not be resolved, but is saved
also to be used in Section
\ref{sec:robustSIFTFeatureMatching}. Finally, in the case where we
have no matches, we can still perform rotations in order to exploit
patterns in indoor environments. For instance, doors, windows,
furniture, etc. all tend to have linear edges that are parallel to the
edges of the surfaces they are on. Similarly, lights, visible interior
walls, etc. which are visible in floor and ceiling images, all tend to
be parallel to exterior walls corresponding to the edges of ceiling
and floor surfaces. Thus, we choose to minimize the angle between
image-based lines and any geometry-based line. We do this using the
RANSAC framework, which successfully ignores outliers
\cite{fischler1981random}.

After these steps are performed, our image projections line up well with our target surface, as shown in Figure \ref{fig:geometryAlignment}. This procedure reconciles both errors in camera poses as well as in geometry, and results in sharp, continuous borders across images, which is crucial when performing occlusion checks.

\subsection{Image Occlusion}
\label{sec:imageOcclusion}
In order to properly texture surfaces such as the one shown in Figure
\ref{fig:occlusion}, it is important to detect when parts of image
projections contain texture belonging to other surfaces and should
therefore not be used. This can be tested by recursively performing
ray-polygon intersection tests between camera locations and other
surfaces in a regularly spaced grid \cite{rayintersection}. Where four
corners of a rectangular region are occluded, texture is
removed. Where no corners are occluded, the recursion stops. Where
there is a mixture of both, the rectangular region is subdivided into
four, and the same process is performed on each. By performing Section
\ref{sec:geometryAlignment}'s alignment procedure before occlusion,
texture belonging to other surfaces is accurately removed, which is
necessary for the next section.

\subsection{2D Feature Alignment}
\label{sec:robustSIFTFeatureMatching}
Our next step is to align images to each other by searching for
corresponding feature points between all pairs of overlapping
images. We use feature alignment rather than pixel or intensity-based
alignment due to the differences in lighting as well as possible
occlusion among our images, both of which feature alignment is less
sensitive to \cite{lowe1999object, mikolajczyk2005performance,
  szeliski2006image}.  We use SiftGPU \cite{siftgpu} for its high
performance on both feature detection as well as pairwise
matching. These matches determine $dx$ and $dy$ distances between each
pair of features for two image projections, though these distances may
not always be the same for different features. Since indoor
environments often contain repetitive features such as floor tiles or
doors, we need to ensure that SIFT-based distances are
reliable. First, we only perform alignment on parts of images that
overlap given the original noisy poses. Second, we discard feature
matches that correspond to an image distance greater than 200 mm from
what the noisy poses estimate. In order to utilize the remaining
feature matches robustly, RANSAC \cite{fischler1981random} is again
used to estimate the optimal $dx_{i,j}$ and $dy_{i,j}$ distances
between two images $i$ and $j$. We use a 5 mm threshold for RANSAC, so
that SIFT matches are labeled as outliers if their distance is not
within 5 mm of the sampled average distance.

We now use the feature-based distances between each pair of images as
well as our previous geometry alignment results to refine all image
positions using a weighted linear least squares solution. The
variables we wish to solve for are the $x_i$ and $y_i$ positions of
images, while equations are the feature-based distances between pairs
of images, images fixed to geometry with 0 or 1 degrees of freedom,
and the original noisy camera poses. The original camera poses are
needed because we sometimes are not able to detect feature matches in
all of our images, or we lack enough geometry alignment results to
generate a single solution. Because we want to minimally use our
original poses, we give them a weighting factor of $0.01$, while all
other equations are weighted at $1$. An example setup for a weighted
linear least squares problem $\textrm{min}_{\vec{\beta}}
||W^\frac{1}{2}(A \vec{\beta} - \vec{\gamma})||_2^2 $ with 3 images is
shown below.

\[
A =
\begin{pmatrix}
  -1 & 1 & 0 & 0 & 0 & 0\\
  0 & 0 & 0 & -1 & 1 & 0\\
  0 & -1 & 1 & 0 & 0 & 0\\
  0 & 0 & 0 & 0 & -1 & 1\\
  0 & -m_2 & 0 & 0 & 1 & 0\\
  1 & 0 & 0 & 0 & 0 & 0\\
  0 & 0 & 0 & 1 & 0 & 0\\
  1 & 0 & 0 & 0 & 0 & 0\\
  0 & 0 & 0 & 1 & 0 & 0\\

\end{pmatrix}\quad
\vec{\beta} =
\begin{pmatrix}
  x_1, \\ x_2, \\ x_3, \\ y_1, \\ y_2, \\ y_3
\end{pmatrix}
\vec{\gamma} =
\begin{pmatrix}
  dx_{1,2}, \\ dy_{1,2}, \\ dx_{2,3}, \\ dy_{2,3}, \\ -m_2gx_2 + gy_2,
  \\ gx_1, \\ gy_1, \\ tx_1, \\ ty_1

\end{pmatrix}
\vec{W} =
\begin{pmatrix}
  1, \\ 1, \\ 1, \\ 1, \\ 1, \\ 1, \\ 1, \\ 0.01, \\ 0.01
\end{pmatrix}
\]

In this scenario, a feature-based distance of $dx_{1,2}$, $dy_{1,2}$
was calculated between images 1 and 2. This corresponds to the first
and second row of $A$, while the third and fourth row of $A$ represent
the same for images 2 and 3. Rows 5 through 7 correspond to results of
the geometry alignment procedure in Section
\ref{sec:geometryAlignment}. Specifically, row 5 corresponds to a
geometry-based constraint of image 2's location to a line of slope
$m_2$, passing through point $gx_2$, $gy_2$, while rows 6 and 7
correspond to a fixed location for image 1 without any degrees of
freedom. Rows 8 and 9 correspond to the original camera pose for image
1, and original poses for images 2 and 3 are omitted for brevity.

Because our problem is linear, it is easily and quickly solved, and
after applying the resulting shifts, our images overlap and match each
other with far greater accuracy. Applying the simple mapping scheme in
Section \ref{sec:tileBasedMapping} to the same wall used in that
section results in Figure \ref{fig:compareAll}(b), which has far fewer
discontinuities, though errors due to lighting differences and
parallax effects are still visible.

\section{Image Compositing}
\label{sec:imageCompositing}
In this section, we revisit the tile-based texturing approach from
Section \ref{sec:simpleTextureMapping}, with an added caching
mechanism to reduce image boundaries. This method works well given all
manner of camera poses and surfaces, but for optimal cases where
images have consistently perpendicular viewing angles, we propose a
superior method that selectively reduces image boundaries.

\subsection{Tile-Mapping with Caching}
\label{sec:mappingWithCaching}
From Section \ref{sec:tileBasedMapping}, we saw that discontinuities
occur where adjacent tiles are textured by different images. Though
Section \ref{sec:2dAlignment}'s image alignment removes many such
discontinuities, there are still cases where seams are visible due to
imprecise matching or other factors such as model-based errors. To
reduce the cases where this happens, it makes sense to take into
account image selections made by neighboring tiles while texture
mapping a given tile. By using the same image across tile boundaries,
we can eliminate a discontinuity altogether. If this is not possible
because a tile is not visible in images used by neighboring tiles,
using similar images across tile boundaries also leads to less
noticeable discontinuities.

Essentially a caching mechanism, we select the best image for a tile
$t$ by searching through two subsets of images for a viable candidate,
before searching through the entire set of valid images. The first
subset of images is the images selected by adjacent tiles that have
already been textured. We must first check which of these images can
map to $t$, and then of those, we make a choice according to the
scoring function in Figure \ref{fig:scoringFunction}. Before reusing
this image, we ensure it meets the criteria $\alpha < 45^\circ$, in
order to ensure a high resolution projection, with $\alpha$ as the
camera angle as shown in Figure \ref{fig:scoringFunction}.

If no satisfactory image is found in the first subset, we check a
second subset of images, consisting of those taken near the ones in
the first subset, both spatially and temporally. These images are not
the same as the ones used for neighboring tiles, but are taken at a
similar location and time, suggesting that their localization and
projection are quite similar, and thus likely matched more cleanly. If
no viable image is found according to the same criteria as before, we
search the entire set of candidate images, selecting based on the same
scoring function from Figure \ref{fig:scoringFunction}.

The result of this caching approach is shown in Figure
\ref{fig:compareAll}(c), where seams are now reduced as compared to
Figure \ref{fig:compareAll}(b). However, some discontinuities are
still present, as visible in the posters on the wall with breaks in
their borders.

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=1in]{projectionWall.pdf}}
  \centering
  \subfloat[][]{\includegraphics[width=1.1in]{projectionCeiling.pdf}}
  \centering
  \subfloat[][]{\includegraphics[width=0.9in]{projectionWallCrop.pdf}}
  \caption{(a) Images for vertical planes are tilted, but their camera
    axes are more or less normal to their respective planes. (b)
    Camera axes for ceiling images are at large angles with respect to
    plane normals. (c) Wall images are cropped to be rectangular.}
  \label{fig:projectionAngles}
\end{figure}

As mentioned earlier, our data comes from a mobile backpack
system. Human operators can not carry the backpack in a perfectly
upright position and are bent forwards at 15 to 20 degrees with
respect to the vertical direction. As a result, cameras facing
sideways are head on with respect to vertical walls, while cameras
oriented towards the top or bottom of the backpack are at an angle
with respect to horizontal floors and ceilings. This is depicted in
Figures \ref{fig:projectionAngles}(a) and
\ref{fig:projectionAngles}(b). These oblique camera angles for
horizontal surfaces translate into textures that span large areas once
projected, as shown in Figure \ref{fig:projectionAngles}(b). Using the
tile-based texture mapping criteria from Figure
\ref{fig:scoringFunction}, such projections have highly varying scores
depending on the location of a tile on the plane. Thus, the tiling
approach in this section is a good choice for texturing floors and
ceilings, as it uses the parts of image projections that maximize
resolution and accuracy for their respective plane locations,
e.g. areas near point A and not near point B, in Figure
\ref{fig:projectionAngles}(b).

\subsection{Shortest Path Texturing}
\label{sec:shortestPath}
For vertical walls, most images are taken from close distances and
head-on angles, resulting in high resolution fronto-parallel
projections. As a result, for each tile on a wall plane, the scoring
function of Figure \ref{fig:scoringFunction} is relatively flat with
respect to candidate images, as they are all more or less head
on. Thus, the scoring function is less significant for walls, and it
is conceivable to use a different texturing strategy to directly
minimize visible seams when texturing them. This is done by choosing
the smallest possible set of images that (a) covers the entire plane
and (b) minimizes the visibility of borders between them. A
straightforward cost function that accomplishes the latter is the sum
of squared differences (SSD) of pixels in overlapping regions between
all pairs of images. Minimizing this cost function encourages image
boundaries to occur either in featureless areas, such as bare walls,
or in areas where images match extremely well.

The first step in this procedure is to obtain a list of images useful
for texturing the wall. A simple way to do this is to use the images
selected by the tiling process in Section
\ref{sec:tileBasedMapping}. Such a list of images is guaranteed to
cover the entire wall, and consists of desired camera poses overall.

In the general case, our problem can be defined as minimally covering
a polygon i.e. the planar surface, using other polygons of arbitrary
geometry i.e. image projections, with the added constraint of
minimizing the cost function between chosen images. Given that
wall-texture candidate images are taken from more or less head-on
angles, and knowing that only minor rotations are made in Section
\ref{sec:2dAlignment}, we can crop our image projections to be
rectangular with minimal texture loss. Furthermore, because our
fisheye camera lenses have full floor-to-ceiling coverage of nearly
all walls, and our backpack operator logically only moves
horizontally, we only need to ensure lateral coverage of our wall
planes. We can thus construct a Directed Acyclic Graph (DAG) from the
images, with edge costs defined by the SSD cost function, and solve a
simple shortest path problem to find an optimal subset of images with
regard to the SSD cost function \cite{dijkstra}.

\begin{figure}
  \centering
  \includegraphics[width=5in]{dagCreation.pdf}
  \caption{DAG construction for the image selection process. \\}
  \label{fig:dagCreation}
\end{figure}

Figure \ref{fig:dagCreation} demonstrates the construction of a DAG
from overlapping images of a hallway wall. Images are sorted by
horizontal location left to right, and become nodes in a
graph. Directed edges are placed in the graph from left to right
between overlapping images. The weights of these edges are determined
by the SSD cost function. Next, we add two artificial nodes, one start
node representing the left border of the plane, and one end node
representing the right border of the plane. The left(right) artificial
node has directed edges with equal and arbitrary cost $C_0$ to(from)
all images that meet the left(right) border of the plane. We now solve
the shortest path problem from the start node to the end node. This
results in a set of images completely covering the plane horizontally,
while minimizing the cost of seams between images.

In very rare cases where the vertical dimension of the plane is not
entirely covered by one or more chosen images, we are left with holes
where no images are selected to texture. Since these holes are rare,
and generally fairly small, we use a greedy approach, repeatedly
filling the hole with images that result in the lowest SSD costs in a
blending region around the hole, as discussed in Section
\ref{sec:blending}. This method is not as optimal as a true
2D-coverage solution would be, but it is a fast approximation, and
adequately handles the few holes we encounter.

With this completed, we have now mapped every location on the plane to
at least one image, and have minimized the number of images, as well
as the discontinuities at their borders. As seen in Figure
\ref{fig:compareAll}(d), this shortest path method has fewer visible
discontinuities than Figure \ref{fig:compareAll}(c) corresponding to
the tile caching approach\footnote{In Figure \ref{fig:compareAll}(d),
  we arbitrarily chose one image for texturing where images overlap,
  as blending will be discussed in section \ref{sec:blending}.}. This
is especially evident when comparing the posters in the images. This
shortest path approach approach directly reduces the cost of each
image boundary, while the tile caching method uses a scoring function
that only approximates this effect. Furthermore, this approach
guarantees the best selection of images to minimize seams, while the
sequential tile caching method may select images early on that turn
out to be poor choices once subsequent tiles have been processed. This
shortest path approach is also far less intensive in terms of memory
usage and runtime, both during texture generation and rendering, as it
does not require discretizing planes or images.

When texturing an entire 3D planar model, we apply the shortest path
method on walls, due to its superior output when provided with head-on
images. Floors and ceilings however, given their many images taken at
oblique angles, are textured using the tile caching method.

\subsection{Blending}
\label{sec:blending}
We now apply a blending procedure to both texturing methods. Although
the image alignment steps and image selection in both methods attempt
to minimize all mismatches between images, there are occasional
unavoidable discontinuities in the final texture due to different
lighting conditions or inaccuracies in model geometry. These can
however be treated and smoothed over by applying alpha blending over
image seams.  Whether the units we are blending are
rectangularly-cropped images or rectangular tiles, we can apply the
same blending procedure, as long as we have a guaranteed overlap
between units to blend over.

For the tile caching method, we can ensure overlap by texturing a
larger tile than needed for display. For example, for a rendered tile
$l_1 \times l_1$, we can associate it with a texture $(l_1 + l_2)
\times (l_1 + l_2)$ in size.  We have found $l_2 = \frac{l_1}{2}$ to
provide a balance between blending and keeping features unblurred. For
the shortest path method, we have already ensured overlap between
images. To enforce consistent blending however, we add a minimum
required overlap distance of 200 mm while solving the shortest path
problem in Section \ref{sec:shortestPath}. Additionally, if images
overlap in a region greater than the overlap distance, we only apply
blending over an area equal to the overlap distance.

After performing linear alpha blending across overlapping regions, our
texture mapping process is complete. Figures \ref{fig:compareAll}(e)
and \ref{fig:compareAll}(f) show the blended versions of Figures
\ref{fig:compareAll}(c) and \ref{fig:compareAll}(d) respectively. The
remaining images in Figure \ref{fig:compareAll} highlight differences
between the two methods, showing that Figure \ref{fig:compareAll}(f)
has the best visual quality and the best texturing approach among the
textures in Figure \ref{fig:compareAll}.

\section{Results and Conclusions}
\label{sec:resultsAndConclusions}
Examples of ceilings and floors textured with the tile caching
approach, and walls textured with the shortest path approach, are
displayed in Figure \ref{fig:results}. These results are as accurate
and clean as the best results we obtained using the two algorithms
mentioned in Section \ref{sec:existingApproaches}. High resolution
colored texture comparisons, as well as video and interactive
walkthroughs of full models are available at
\footnote{\url{http://www.eecs.berkeley.edu/~pcheng/indoormapping}}.

As mentioned earlier, our approach is quite efficient. The top wall in
Figure \ref{fig:results}(a) was generated with 7543 $\times$ 776
pixels, and spans a 40-meter long wall. Given 41000 input images of
the entire dataset, a 2.8GHz dual-core consumer-grade laptop takes
approximately a minute to pick 36 candidate images, followed by under
a minute to perform both image alignment and the shortest path
texturing method, though over 75\% of that time is spent calculating
SIFT matches within the SiftGPU framework, which could feasibly be
split into a separate preprocessing step. While not real-time, the
process is capable of generating quick updates after changes in
various parameters or modifications to input data, and if integrated
directly into a modeling system, could provide live visualization as
data is collected. Our full models consist of our input model file,
textures that we generate, and a mapping of image points to model
vertices. The models shown in Figure \ref{fig:results} are roughly 20
MB in size, and are visualized using the OpenSceneGraph toolkit
\cite{openscenegraph}, which allows for export to many common model
formats, as well as efficient visualization, even in web browsers or
mobile devices.

In this paper, we have developed an approach to texture mapping models
with noisy camera localization data. We are able to refine image
locations based on geometry references and feature matching, and
robustly handle outliers. Using the tile-based mapping approach, we
can texture both simple rectangular walls as well as complex floor and
ceiling geometry. We also implemented a shortest path texturing method
that produces seamless textures on planes where multiple head-on
images are available. Both of these approaches are highly modular, and
easily tunable for similar systems across multiple environments.

While our method works well on our datasets, it is likely to fail in
areas where 3D error is much higher. A technique for resolving camera
error in 3D that would be a logical progression of our approach is to
perform 3D matching between image lines and geometry in 3D, which can
be done reasonably efficiently \cite{linebased,
  rectangularstructures}. Using linear features on top of our SIFT
features is also likely to result in improved matches, as indoor
scenes often have long, unbroken lines spanning multiple images
\cite{linearposeestimation}. Finally, our blending procedure is very
basic, and applying more sophisticated methods of blending as well as
normalization would benefit our final visual quality, and more
robustly handle motion-based or parallax errors.

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=3in]{finalfloors.jpg}} ~~~~~~~~
  \centering
  \subfloat[][]{\includegraphics[width=3in]{finalceilings.jpg}}

  \centering \subfloat[][]{
    \includegraphics[height=2in, width=3in]{floorcropped.jpg} ~~~~~~~~
    \includegraphics[width=3in]{pier15floor.jpg}
  }

  \centering{
    \includegraphics[width=1.9in]{fullmodel.png}
    \includegraphics[width=1.9in]{threestoryfull2.png}
    \includegraphics[width=1.9in]{pier15.png} \\
    \includegraphics[width=6in]{threestoryfull.png}
  }

  \centering \subfloat[][]{ }
  \caption{Examples of our final texture mapping output for (a) walls,
    (b) ceilings, (c) floors, (d) full models.}
  \label{fig:results}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%% References %%%%%

\bibliography{report} %>>>> bibliography data in report.bib
\bibliographystyle{spiebib} %>>>> makes bibtex use spiebib.bst

\end{document}