### 2013-05-02 masters/body/assets/3dimpvt/template/egpaper_for_review.tex

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

% \cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{****} % *** Enter the 3DIMPVT Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
\ifcvprfinal\pagestyle{empty}\fi
\begin{document}

%%%%%%%%% TITLE
\title{\LaTeX\ Author Guidelines for 3DIMPVT Proceedings}

\author{First Author\\
  Institution1\\
  Institution1 address\\
  {\tt\small firstauthor@i1.org}
  % For a paper whose authors are all at the same institution, omit
  % the following lines up until the closing ``}''.  Additional
  % authors and addresses can be added with ``\and'', just like the
  % second author.  To save space, use either the email address or
  % home page, not both
  \and
  Second Author\\
  Institution2\\
  First line of institution2 address\\
  {\small\url{http://www.author.org/~second}} }

\maketitle
% \thispagestyle{empty}

%%%%%%%%% ABSTRACT
\begin{abstract}
  abcd The ABSTRACT is to be in fully-justified italicized text, at the top
  of the left-hand column, below the author and affiliation
  information. Use the word ``Abstract'' as the title, in 12-point
  Times, boldface type, centered relative to the column, initially
  capitalized. The abstract is to be in 10-point, single-spaced type.
  Leave two blank lines after the Abstract, then begin the main text.
  Look at previous 3DIMPVT/3DIM/3DPVT abstracts to get a feel for
  style and length.
\end{abstract}

%%%%%%%%% BODY TEXT
\section{Introduction}

Please follow the steps outlined below when submitting your manuscript
to the IEEE Computer Society Press.  This style guide now has several
important modifications (for example, you are no longer warned against
the use of sticky tape to attach your artwork to the paper), so all
authors should read this new version.

% -------------------------------------------------------------------------
\subsection{Language}

All manuscripts must be in English.

\subsection{Dual submission}

By submitting a manuscript to 3DIMPVT, the authors assert that it has
not been previously published in substantially similar
form. Furthermore, no paper which contains significant overlap with
the contributions of this paper either has been or will be submitted
during the 3DIMPVT 2012 review period to {\bf either a journal} or any
conference or any workshop.  {\bf Papers violating this condition will
  be rejected}.

If there are papers that may appear to the reviewers to violate this
condition, then it is your responsibility to: (1)~cite these papers
(preserving anonymity as described in Section 1.6 below), (2)~argue in
the body of your paper why your 3DIMPVT paper is non-trivially
different from these concurrent submissions, and (3)~include
anonymized versions of those papers in the supplemental material.

\subsection{Paper length}
3DIMPVT papers may be between 6 pages and 8 pages.  Overlength papers
will simply not be reviewed.  This includes papers where the margins
and formatting are deemed to have been significantly altered from
those laid down by this style guide.  Note that this \LaTeX\ guide
already sets figure captions and references in a smaller font.  The
reason such papers will not be reviewed is that there is no provision
for supervised revisions of manuscripts.  The reviewing process cannot
determine the suitability of the paper for presentation in eight pages
if it is reviewed in eleven.

% -------------------------------------------------------------------------
\subsection{The ruler}
The \LaTeX\ style defines a printed ruler which should be present in
the version submitted for review.  The ruler is provided in order that
reviewers may comment on particular lines in the paper without
circumlocution.  If you are preparing a document using a non-\LaTeX\
document preparation system, please arrange for an equivalent ruler to
appear on the final output pages.  The presence or absence of the
ruler should not change the appearance of any other content on the
page.  The camera ready copy should not contain a ruler. (\LaTeX\
users may uncomment the \verb'\cvprfinalcopy' command in the document
preamble.)  Reviewers: note that the ruler measurements do not align
well with lines in the paper --- this turns out to be very difficult
to do well when the paper contains many figures and equations, and,
when done, looks ugly.  Just use fractional references (e.g.\ this
line is $097.15$), although in most cases one would expect that the
approximate location will be adequate.

\subsection{Mathematics}

Please number all of your sections and displayed equations.  It is
important for readers to be able to refer to any particular equation.
Just because you didn't refer to it in the text doesn't mean some
future reader might not need to refer to it.  It is cumbersome to have
to use circumlocutions like ``the equation second from the top of page
3 column 1''.  (Note that the ruler will not be present in the final
copy, so is not an alternative to equation numbers).  All authors will
benefit from reading Mermin's description of how to write mathematics:
\url{http://www.cvpr.org/doc/mermin.pdf}.

\subsection{Blind review}

Many authors misunderstand the concept of anonymizing for blind
review.  Blind review does not mean that one must remove citations to
one's own work---in fact it is often impossible to review a paper
unless the previous citations are known and available.

Blind review means that you do not use the words ``my'' or ``our''
when citing previous work.  That is all.  (But see below for
techreports)

Saying ``this builds on the work of Lucy Smith [1]'' does not say that
you are Lucy Smith, it says that you are building on her work.  If you
are Smith and Jones, do not say ``as we show in [7]'', say ``as Smith
and Jones show in [7]'' and at the end of the paper, include reference
7 as you would any other cited work.

An example of a bad paper just asking to be rejected:
\begin{quote}
  \begin{center}
    An analysis of the frobnicatable foo filter.
  \end{center}

  In this paper we present a performance analysis of our previous
  paper [1], and show it to be inferior to all previously known
  methods.  Why the previous paper was accepted without this analysis
  is beyond me.

  [1] Removed for blind review
\end{quote}

An example of an acceptable paper:

\begin{quote}
  \begin{center}
    An analysis of the frobnicatable foo filter.
  \end{center}

  In this paper we present a performance analysis of the paper of
  Smith \etal [1], and show it to be inferior to all previously known
  methods.  Why the previous paper was accepted without this analysis
  is beyond me.

  [1] Smith, L and Jones, C. ``The frobnicatable foo filter, a
  fundamental contribution to human knowledge''.  Nature 381(12),
  1-213.
\end{quote}

If you are making a submission to another conference at the same time,
which covers similar or overlapping material, you may need to refer to
that submission in order to explain the differences, just as you would
if you had previously published related work.  In such cases, include
the anonymized parallel submission~\cite{Authors12} as additional
material and cite it as
\begin{quote} [1] Authors. ``The frobnicatable foo filter'', F\&G 2012
  Submission ID 324, Supplied as additional material {\tt fg324.pdf}.
\end{quote}

Finally, you may feel you need to tell the reader that more details
can be found elsewhere, and refer them to a technical report.  For
conference submissions, the paper must stand on its own, and not {\em
  require} the reviewer to go to a techreport for further details.
Thus, you may say in the body of the paper ``further details may be
found in~\cite{Authors12b}''.  Then submit the techreport as
additional material.  Again, you may not assume the reviewers will
read this material.

Sometimes your paper is about a problem which you tested using a tool
which is widely known to be restricted to a single institution.  For
example, let's say it's 1969, you have solved a key problem on the
Apollo lander, and you believe that the CVPR70 audience would like to
hear about your solution.  The work is a development of your
celebrated 1968 paper entitled ``Zero-g frobnication: How being the
only people in the world with access to the Apollo lander source code
makes us a wow at parties'', by Zeus \etal.

You can handle this paper like any other.  Don't write ``We show how
to improve our previous work [Anonymous, 1968].  This time we tested
the algorithm on a lunar lander [name of lander removed for blind
review]''.  That would be silly, and would immediately identify the
authors. Instead write the following:
\begin{quotation}
  \noindent
  We describe a system for zero-g frobnication.  This system is new
  because it handles the following cases: A, B.  Previous systems
  [Zeus et al. 1968] didn't handle case B properly.  Ours handles it
  by including a foo term in the bar integral.

  ...

  The proposed system was integrated with the Apollo lunar lander, and
  went all the way to the moon, don't you know.  It displayed the
  following behaviours which show how well we solved cases A and B:
  ...
\end{quotation}
As you can see, the above text follows standard scientific convention,
reads better than the first version, and does not explicitly name you
as the authors.  A reviewer might think it likely that the new paper
was written by Zeus \etal, but cannot make any decision based on that
guess.  He or she would have to be sure that no other authors could
have been contracted to solve problem B.

FAQ: Are acknowledgements OK?  No.  Leave them for the final copy.

\begin{figure}[t]
  \begin{center}
    \fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
    % \includegraphics[width=0.8\linewidth]{egfigure.eps}
  \end{center}
  \caption{Example of caption.  It is set in Roman so that mathematics
    (always set in Roman: $B \sin A = A \sin B$) may be included
    without an ugly clash.}
  \label{fig:long}
  \label{fig:onecol}
\end{figure}

\subsection{Miscellaneous}

\noindent
Compare the following:\\
\begin{tabular}{ll}
  \verb'$conf_a$' &  $conf_a$ \\
  \verb'$\mathit{conf}_a$' & $\mathit{conf}_a$
\end{tabular}\\
See The \TeX book, p165.

The space after \eg, meaning ``for example'', should not be a
sentence-ending space. So \eg is correct, {\em e.g.} is not.  The
provided \verb'\eg' macro takes care of this.

When citing a multi-author paper, you may save space by using ``et
alia'', shortened to ``\etal'' (not ``{\em et.\ al.}'' as ``{\em et}''
is a complete word.)  However, use it only when there are three or
more authors.  Thus, the following is correct: `` Frobnication has
been trendy lately.  It was introduced by Alpher~\cite{Alpher02}, and
subsequently developed by Alpher and
Fotheringham-Smythe~\cite{Alpher03}, and Alpher
\etal~\cite{Alpher04}.''

This is incorrect: ``... subsequently developed by Alpher
\etal~\cite{Alpher03} ...''  because reference~\cite{Alpher03} has
just two authors.  If you use the \verb'\etal' macro provided, then
you need not worry about double periods when used at the end of a
sentence as in Alpher \etal.

For this citation style, keep multiple citations in numerical (not
chronological) order, so prefer \cite{Alpher03,Alpher02,Authors12} to
\cite{Alpher02,Alpher03,Authors12}.

\begin{figure*}
  \begin{center}
    \fbox{\rule{0pt}{2in} \rule{.9\linewidth}{0pt}}
  \end{center}
  \caption{Example of a short caption, which should be centered.}
  \label{fig:short}
\end{figure*}

% ------------------------------------------------------------------------
\section{Formatting your paper}

All text must be in a two-column format. The total allowable width of
the text area is $6\frac78$ inches (17.5 cm) wide by $8\frac78$ inches
(22.54 cm) high. Columns are to be $3\frac14$ inches (8.25 cm) wide,
with a $\frac{5}{16}$ inch (0.8 cm) space between them. The main title
(on the first page) should begin 1.0 inch (2.54 cm) from the top edge
of the page. The second and following pages should begin 1.0 inch
(2.54 cm) from the top edge. On all pages, the bottom margin should be
1-1/8 inches (2.86 cm) from the bottom edge of the page for $8.5
\times 11$-inch paper; for A4 paper, approximately 1-5/8 inches (4.13
cm) from the bottom edge of the page.

% -------------------------------------------------------------------------
\subsection{Margins and page numbering}

All printed material, including text, illustrations, and charts, must
be kept within a print area 6-7/8 inches (17.5 cm) wide by 8-7/8
inches (22.54 cm) high.

% -------------------------------------------------------------------------
\subsection{Type-style and fonts}

Wherever Times is specified, Times Roman may also be used. If neither
is available on your word processor, please use the font closest in
appearance to Times to which you have access.

MAIN TITLE. Center the title 1-3/8 inches (3.49 cm) from the top edge
of the first page. The title should be in Times 14-point, boldface
type.  Capitalize the first letter of nouns, pronouns, verbs,
adjectives, and adverbs; do not capitalize articles, coordinate
conjunctions, or prepositions (unless the title begins with such a
word). Leave two blank lines after the title.

AUTHOR NAME(s) and AFFILIATION(s) are to be centered beneath the title
and printed in Times 12-point, non-boldface type. This information is
to be followed by two blank lines.

The ABSTRACT and MAIN TEXT are to be in a two-column format.

MAIN TEXT. Type main text in 10-point Times, single-spaced. Do NOT use
double-spacing. All paragraphs should be indented 1 pica (approx. 1/6
inch or 0.422 cm). Make sure your text is fully justified---that is,
flush left and flush right. Please do not place any additional blank
lines between paragraphs.

Figure and table captions should be 9-point Roman type as in
Figures~\ref{fig:onecol} and~\ref{fig:short}.  Short captions should
be centred.

\noindent Callouts should be 9-point Helvetica, non-boldface type.
Initially capitalize only the first word of section titles and first-,
second-, and third-order headings.

FIRST-ORDER HEADINGS. (For example, {\large \bf 1. Introduction})
should be Times 12-point boldface, initially capitalized, flush left,
with one blank line before, and one blank line after.

SECOND-ORDER HEADINGS. (For example, { \bf 1.1. Database elements})
should be Times 11-point boldface, initially capitalized, flush left,
with one blank line before, and one after. If you require a
third-order heading (we discourage it), use 10-point Times, boldface,
initially capitalized, flush left, preceded by one blank line,
followed by a period and your text on the same line.

% -------------------------------------------------------------------------
\subsection{Footnotes}

Please use footnotes\footnote {This is what a footnote looks like.  It
  often distracts the reader from the main flow of the argument.}
sparingly.  Indeed, try to avoid footnotes altogether and include
necessary peripheral observations in the text (within parentheses, if
you prefer, as in this sentence).  If you wish to use a footnote,
place it at the bottom of the column on the page on which it is
referenced. Use Times 8-point type, single-spaced.

% -------------------------------------------------------------------------
\subsection{References}

List and number all bibliographical references in 9-point Times,
single-spaced, at the end of your paper. When referenced in the text,
enclose the citation number in square brackets, for
example~\cite{Authors12}.  Where appropriate, include the name(s) of
editors of referenced books.

\begin{table}
  \begin{center}
    \begin{tabular}{|l|c|}
      \hline
      Method & Frobnability \\
      \hline\hline
      Theirs & Frumpy \\
      Yours & Frobbly \\
      Ours & Makes one's heart Frob\\
      \hline
    \end{tabular}
  \end{center}
  \caption{Results.   Ours is better.}
\end{table}

% -------------------------------------------------------------------------
\subsection{Illustrations, graphs, and photographs}

All graphics should be centered.  Please ensure that any point you
wish to make is resolvable in a printed copy of the paper.  Resize
fonts in figures to match the font in the body text, and choose line
widths which render effectively in print.  Many readers (and
reviewers), even of an electronic copy, will choose to print your
paper in order to read it.  You cannot insist that they do otherwise,
and therefore must not assume that they can zoom in to see tiny
details on a graphic.

When placing figures in \LaTeX, it's almost always best to use
\verb+\includegraphics+, and to specify the figure width as a multiple
of the line width as in the example below {\small\begin{verbatim}
    \usepackage[dvips]{graphicx} ...
   \includegraphics[width=0.8\linewidth]
                   {myfile.eps}
\end{verbatim}
}

% -------------------------------------------------------------------------
\subsection{Color}

Color is valuable, and will be visible to readers of the electronic
copy.  However ensure that, when printed on a monochrome printer, no
important information is lost by the conversion to grayscale.

% ------------------------------------------------------------------------
\section{Final copy}

You must include your signed IEEE copyright release form when you
submit your finished paper. We MUST have this form before your paper
can be published in the proceedings.

% Please direct any questions to the production editor in charge of
% these proceedings at the IEEE Computer Society Press: Phone (714)
% 821-8380, or Fax (714) 761-1784.

{\small \bibliographystyle{ieee} \bibliography{egbib} }

\end{document}