### 2013-12-29 acivs/acivs.tex
cleanup

% This is LLNCS.DEM the demonstration file of
% the LaTeX macro package from Springer-Verlag
% for Lecture Notes in Computer Science,
% version 2.4 for LaTeX2e as of 16. April 2010
%
\documentclass{llncs}
\usepackage{cite}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{float}
\usepackage{multirow}
\title{Texture mapping 3D planar models of indoor environments with noisy camera poses}

%>>>> The author is responsible for formatting the 
%  author list and their institutions.  Use  \skiplinehalf 
%  to separate author list from addresses and between each address.
%  The correspondence between each author and his/her address
%  can be indicated with a superscript in italics, 
%  which is easily obtained with \supit{}.

\author{Peter Cheng \and Michael Anderson \and Stewart He \and Avideh Zakhor}
\institute{University of California, Berkeley\\ Berkeley CA 94720, USA}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
%>>>> uncomment following for page numbers
% \pagestyle{plain}    
%>>>> uncomment following to start page numbering at 301 
%\setcounter{page}{301}

\begin{document}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\begin{abstract}
  Automated 3D modeling of building interiors is used in applications
  such as virtual reality and environment mapping. Texturing these
  models allows for photo-realistic visualizations of the data
  collected by such modeling systems. While data acquisition times for
  mobile mapping systems are considerably shorter than for static
  ones, their recovered camera poses often suffer from inaccuracies,
  resulting in visible discontinuities when successive images are
  projected onto a surface for texturing. We present a method for
  texture mapping that starts by selecting images whose camera poses
  are well-aligned in two dimensions. We then align images to geometry
  as well as to each other, producing visually consistent textures
  even in the presence of inaccurate surface geometry and noisy camera
  poses. Images are then composited into a final texture mosaic. The
  effectiveness of the proposed method is demonstrated on a number of
  different indoor environments.

  \keywords{Texture Mapping, Reconstruction, Image Stitching, Mosaicing}
\end{abstract}

% >>>> Include a list of keywords after the abstract

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:introduction} % \label{} allows reference to this section
Three-dimensional modeling of indoor environments has a variety of
applications such as training and simulation for disaster management,
virtual heritage conservation, and mapping of hazardous sites. Manual
construction of these digital models can be time consuming, and as
such, automated 3D site modeling has garnered much interest in recent
years.

The first step in automated 3D modeling is the physical scanning of
the environment's geometry. An indoor modeling system must be able to
recover its pose within an environment while simultaneously
reconstructing the 3D structure of the environment itself
\cite{chen2010indoor, hz, kua2012loopclosure, liu2010indoor}. This is known as the simultaneous localization and
mapping (SLAM) problem, and is generally solved by taking readings
from laser range scanners, cameras, and inertial measurement units
(IMUs) at multiple locations within the environment. Mounting such
devices on a platform carried by an ambulatory human provides unique
advantages over vehicular-based systems on wheels in terms of agility
and portability, but can also result in larger localization error
\cite{liu2010indoor}. As a result, common methods for texture mapping
generally produce poor results.

In this paper, we present an approach to texture mapping 3D models of
indoor environments made of planar surfaces in the presence of
uncertainty and noise in camera poses. In particular, we consider data
obtained from a human-operated backpack system with a number of laser
range scanners as well as 2 cameras facing left and right, each
equipped with fisheye lenses reaching an approximately 180$^{\circ}$
field of view and taking photos at a rate of 5 Hz
\cite{liu2010indoor}. Applying multiple localization and loop-closure
algorithms on the raw data collected by the onboard sensors
\cite{chen2010indoor, kua2012loopclosure, liu2010indoor}, the backpack
is localized\footnote{In this paper, we use the terms localization and
  pose recovery interchangeably, in that they both refer to recovering
  position and orientation.}  over its data collection period. This
involves recovering the 6 degrees of freedom for the backpack itself
as well as the cameras rigidly mounted on it. Once this is complete,
the data from the laser range scanners is used to generate a 3D point
cloud of the surrounding environment, from which a 3D planar model is
created \cite{sanchez2012point}. This model, consisting of 2D
polygonal planes in 3D space, along with the set of images captured by
the backpack's cameras and their noisy 3D poses, can be considered the
input to our texture mapping problem.

\begin{figure}
  \centering
  \includegraphics[width=4.5in]{flowchart.jpg}
  \caption{The proposed texture mapping procedure\\}
  \label{fig:flowchart}
\end{figure}

The overall block diagram for the proposed texture mapping procedure
is shown in Figure \ref{fig:flowchart}, where the number attached
attached to each box indicates the section in which the concept in the
box is explained in this paper. We texture map each planar surface
independently and in parallel. For each surface, we begin by selecting
a set of images that spans the entire surface with high resolution
imagery. We then use our noisy camera poses to project these selected
images onto the surface. These projections are then refined in 2D, in
order to maximally align them with the surface's geometry, as well as
to each other, allowing us to handle both errors in geometry as well
as camera poses. For surfaces and camera poses at arbitrary locations
or orientations, generally ceilings and floors, we propose a
tile-based approach for sampling high-resolution portions of images
and compositing them into a texture. In cases where cameras have
consistently perpendicular viewing angles to the surfaces under
consideration, generally walls, we demonstrate a superior approach
that leads to more seamless textures.

The remainder of the paper is organized as follows. Section
\ref{sec:existingApproaches} covers existing approaches to image
stitching, and their performance on our datasets. Section
\ref{sec:simpleTextureMapping} explains how to downsample the set of
available images by selecting those with the best orientation and
distance from surfaces under consideration. Section
\ref{sec:2dAlignment} contains our approach towards 2D image
alignment, followed by Section \ref{sec:imageCompositing}, which
describes two methods of selecting and compositing images to create
the final texture. Sections \ref{sec:results} and \ref{sec:conclusion}
contain results and conclusions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Image Selection}
\label{sec:simpleTextureMapping}

% The geometry of the texture mapping process for a planar surface is
% shown in Figure \ref{fig:projection}. Given a set of $M$ images to
% texture a target plane, camera matrix $P_i$ for the $ith$ image
% transforms a 3D point in the world coordinate system to a 2D point or
% pixel in image $i$'s coordinates. A camera matrix $P_i$ is composed of
% the camera's intrinsic parameters, containing focal length and image
% center, as well as extrinsic parameters which specify the rotation and
% translation of the camera's position in 3D world coordinates at the
% time that image $i$ is taken. These extrinsic parameters are
% determined by the backpack hardware and the corresponding localization
% algorithms \cite{chen2010indoor, liu2010indoor, kua2012loopclosure}
% and are noisy.

Since the backpack system takes pictures at a rate of 5 Hz, hundreds
of images are available for texturing each surface in the 3D model. To
obtain a selection of images to work with, we discretize each surface
into small square tiles, and for each tile, gather the images with a
clear line of sight to it. We then find the image that maximizes the
scoring function $\frac{1}{d} (-1 \cdot \vec{c}) \cdot \vec{n}$ for
each tile, as shown in Figure \ref{fig:scoringFunction}, where $d$ is
the distance between the centers of a camera and a tile, and $\vec{n}$
and $\vec{c}$ are the directions of the plane's normal and the camera
axis respectively. In Figure \ref{fig:compareAll}(a), an image has
been textured using the images selected for each tile. While images
with good viewing angles and high resolution have been used, there are
a significant amount of discontinuities, though many appear
reconcilable with 2D transformations.

By aggregating the images selected for tiles on each surface, we
obtain a collection of images that can be used as a starting point for
the proposed texturing procedure, as explained in Section
\ref{sec:2dAlignment}. For our datasets, this process generally
selects around 10\% of all the images that could be used for texturing
the surface. This roughly 10:1 subsampling removes images with poor
viewing angles, ensuring that the most promising images are considered
in the following steps.

\begin{figure}
  \begin{minipage}[b]{0.45\linewidth}
    \centering
    \includegraphics[height=1.2in,width=2in]{Projection.pdf}
    \caption{Surfaces to be textured are specified in 3D space by
      corners $C_i$. Images are related to each surface through the
      camera matrices $P_{1..m}$. }
    \label{fig:projection}
  \end{minipage}
  \hspace{0.5cm}
  \begin{minipage}[b]{0.45\linewidth}
    \centering
    \includegraphics[height=1.2in, width=2in]{scoringFunction.jpg}
    \caption{We minimize camera angle $\alpha$ and distance $d$ by
      maximizing the scoring function $\frac{1}{d} (-1 \cdot \vec{c})
      \cdot \vec{n}$}
    \label{fig:scoringFunction}
  \end{minipage}
\end{figure}

\begin{figure}
  \centering \subfloat[][]{\includegraphics[width=2.4in,
    height=0.8in]{wall2_naive.jpg}}
  \subfloat[][]{\includegraphics[width=2.4in,
    height=0.8in]{wall2_naive_shift.jpg}}

  \centering \subfloat[][]{\includegraphics[width=2.4in,
    height=0.8in]{wall2_cache_shift.jpg}}
  \subfloat[][]{\includegraphics[width=2.4in,
    height=0.8in]{wall2_shortest_shift.pdf}}

  \centering \subfloat[][]{\includegraphics[width=2.4in,
    height=0.8in]{wall2_cache_shift_blend.pdf}}
  \subfloat[][]{\includegraphics[width=2.4in,
    height=0.8in]{wall2_shortest_shift_blend.pdf}}

  \centering \subfloat[][]{\includegraphics[width=2.4in,
    height=2.2in]{wall2_cache_comparison.png}}
  \subfloat[][]{\includegraphics[width=2.4in,
    height=2.2in]{wall2_shortest_comparison.png}}

  \caption{(a) Tile-based texturing; (b) Tile-based texturing after
    image alignment; (c) Tile-based texturing after image alignment
    with caching; (d) Shortest path texturing after image alignment;
    (e,f) Blending applied to (c) and (d); (g,h) Zoomed in views of
    discontinuities in (e) vs. in (f).}
  \label{fig:compareAll}
\end{figure}

\section{2D Image Alignment}
\label{sec:2dAlignment}
% In this section, we describe our method for efficient and robust image
% alignment. Rather than register all of our images in 3D, as many
% state-of-the-art techniques for image stitching do, we instead select
% a subset of images to be aligned in 2D. This subset corresponds to the
% images selected by the image selection procedure described in Section
% \ref{sec:simpleTextureMapping}.

% Applying 2D alignments to this set of images works well for a number
% of reasons. First, the nature of our input data and the selected
% images is such that localization error chiefly occurs in two
% dimensions, which correspond to the plane of the surface being
% projected onto. This is because the backpack operator, during data
% acquisition, makes efforts to walk within a few meters of, and
% parallel to all walls being scanned. As a result, the translational
% error of camera poses is quite minimal in the dimension perpendicular
% to the wall surface being textured. Furthermore, because our set of
% images has minimal distance from its corresponding wall surface,
% rotational errors are less pronounced, except for rotation around the
% axis perpendicular to the surface being textured. Therefore, most
% translation and rotational errors occur in the 2 dimensions of the
% wall surface's plane. This of course does not apply for ceilings and
% floors; however, the lack of features on most ceilings and floors
% means that 3D misalignment is not likely to visually make a
% difference, beyond what can be done in 2D.

Our proposed 2D alignment procedure consists of three parts, as shown
in the diagram in Figure \ref{fig:flowchart}. First, images are
projected onto the surface and lines within these projected images are
detected. Images are then transformed such that these lines match
geometric lines composing the boundaries of the surface being
textured. Second, occlusion checks are performed to remove invalid
parts of each image for the target surface. Third, SIFT feature
matches are detected between pairs of images, and a weighted linear
least squares problem is solved in order to maximize all image and
geometry-based alignments. Each step will now be explained in detail.

\subsection{Geometry-based Alignment}
\label{sec:geometryAlignment}

\begin{figure}
  \centering
  \begin{tabular}{cc}
    \subfloat[][]{\raisebox{-0.8in}{\includegraphics[trim=0cm 2cm 0cm 0cm,
        clip=true, width=2.4in]{matchLines.jpg}}} &
    \begin{tabular}{c}
      \subfloat[][]{\includegraphics[width=2.4in]{allunshifted.jpg}} \\
      \subfloat[][]{\includegraphics[width=2.4in]{allshifted.jpg}}
    \end{tabular}
  \end{tabular}
  \caption{Images projected onto a ceiling surface, where
    geometry-based lines corresponding to the ceiling's boundary are
    shown in red. Image-based lines detected by Hough transform in the
    image projections are shown in blue; (a) examples of matching
    lines in cases with $\geq$ 2 line pairs, 1 line pair, and zero
    line pairs, from top to bottom; (b) images projected with their
    original noisy camera poses; (c) after image alignment to maximize
    line matches between images and geometry; }
  \label{fig:geometryAlignment}
\end{figure}

To align images to geometry, line segments are detected in images via
Hough transforms. We also gather a set of geometry-based lines, which
correspond to lines comprising the target surface's border, as well as
lines formed where other surfaces intersect the target surface. An
example of these lines is shown in red for a ceiling surface in Figure
\ref{fig:geometryAlignment}(a). Ideally, for perfect camera poses and
surface geometry, the lines in images corresponding to corners between
surfaces should match up exactly with corners in the 3D model. By
inducing such lines to match, we fit camera poses more accurately to
the surface, and therefore to each other as well.

First, pairs of image-based and geometry-based line segments within a
distance and angular difference threshold of each other are
collected. We have found a distance threshold of 250 mm and an
orientation difference threshold of $10^\circ$ to work well for our
datasets. Each such pair signifies a rotation and translation with one
degree of freedom that matches an image line to a geometry line. Pairs
are then binned by rotation in $1^\circ$ increments, which reveals the
rotation that maximizes line matches. From the pairs with that
corresponding rotation, the two longest noncollinear image-based lines
are used to compute a fixed translation. An example of this is case
(1) of Figure \ref{fig:geometryAlignment}(c). If a rotation is
calculated, but a fixed translation cannot be obtained, the pair with
the longest image-based line is selected, and the rotation and the
minimal translation to match its lines are applied. This corresponds
to case (2) of Figure \ref{fig:geometryAlignment}(c). Finally, if no
line pairs were detected, images are rotated using a RANSAC
\cite{fischler1981random} approach to maximize parallel image and
geometry lines, without regard for translational distance. This
corresponds to case (3) of Figure \ref{geometryAlignment}(c). The
results of this procedure are recorded, as the translational degrees
of freedom for each image are used in Section
\ref{sec:robustSIFTFeatureMatching}.

After this step, image projections line up well with target surfaces,
as shown in Figure \ref{fig:geometryAlignment}(b), which is
considerably more aligned than Figure
\ref{fig:geometryAlignment}(a). This procedure reconciles both errors
in camera poses as well as in geometry, and results in sharp,
continuous borders across images, which is crucial when checking for
occlusion.

\subsection{Image Occlusion}
\label{sec:imageOcclusion}

\begin{figure}
  \centering
  \begin{tabular}{cc}
    \subfloat[][]{\raisebox{-0.8in}{\includegraphics[width=2.4in]{occlusiondiagram.png}}} &

    \begin{tabular}{c}
      \subfloat[][]{\includegraphics[trim=0cm 0cm 1.3cm 0cm,
        clip=true, width=2.4in,
        height=0.6in]{occlusionimagebad.png}} \\
      \subfloat[][]{\includegraphics[trim=1.3cm 0cm 0cm 0cm,
        clip=true, width=2.4in,
        height=0.6in]{occlusionimagegood.png}}
    \end{tabular}
  \end{tabular}
\caption{(a) The image from the camera in this diagram contains
  texture that belongs to the gray occluding surface, which should not
  be projected onto the orange target surface; (b) without geometry
  alignment, texture to the left of the red line would be removed,
  which would leave some erroneous texture projected onto our target
  surface; (c) after geometry alignment, the image is shifted,
  resulting in the correct amount of texture being removed.}
\label{fig:occlusion}
\end{figure}

In order to correctly texture surfaces, it is important to detect and
remove parts of image projections containing texture for occluding
surfaces. For instance, in Figure \ref{fig:occlusion}(a), an image
used to texture the orange target surface also contains part of a gray
occluding surface. We remove this incorrect texture by recursively
performing ray-polygon intersection tests between the ray from the
camera location to locations on the target surface
\cite{rayintersection}. These intersection tests are performed at the
corners of a grid overlaid upon the target surface. Where all four
corners of a grid section are occluded, texture is removed. Where one
or more corners are occluded, the grid is subdivided into four, and
the process repeats. Occlusion checking works entirely with geometry,
so by ensuring that images match geometry using
\ref{sec:geometryAlignment}'s alignment procedure, texture belonging
to other surfaces is accurately removed, as seen in Figure
\ref{fig:occlusion}(b).

\subsection{2D Feature Alignment}
\label{sec:robustSIFTFeatureMatching}
The next step is to align images to each other by detecting feature
point matches between all pairs of overlapping images. We use feature
alignment rather than pixel or intensity-based alignment due to the
differences in lighting as well as possible occlusion among images,
both of which feature alignment is less sensitive to
\cite{lowe1999object, mikolajczyk2005performance,
  szeliski2006image}. We use SiftGPU \cite{siftgpu} for its high
performance on both feature detection as well as pairwise
matching. Detected matches determine $dx$ and $dy$ distances between
each pair of features for two image projections, though distances
often are not the same for different features. Since indoor
environments often contain repetitive features such as floor tiles or
doors, we need to ensure that SIFT-based distances are
reliable. First, we only detect feature matches in the parts of images
that overlap given the original noisy poses. Second, feature matches
that correspond to an image distance greater than 200 mm from what the
noisy poses estimate are discarded. In order to utilize the remaining
feature matches robustly, RANSAC \cite{fischler1981random} is used to
estimate the optimal $dx_{i,j}$ and $dy_{i,j}$ distances between two
images $i$ and $j$.

We now use the feature-based distances between pairs of images as well
as geometry alignment results from Section \ref{sec:geometryAlignment}
to refine all image positions using a weighted linear least squares
approach. An example setup for a weighted linear least squares problem
$\textrm{min}_{\vec{\beta}} ||W^\frac{1}{2}(A \vec{\beta} -
\vec{\gamma})||_2^2 $ with 3 images is as follows.

\[
A =
\begin{pmatrix}
  -1 & 1 & 0 & 0 & 0 & 0\\
  0 & 0 & 0 & -1 & 1 & 0\\
  0 & -1 & 1 & 0 & 0 & 0\\
  0 & 0 & 0 & 0 & -1 & 1\\
  0 & -m_2 & 0 & 0 & 1 & 0\\
  1 & 0 & 0 & 0 & 0 & 0\\
  0 & 0 & 0 & 1 & 0 & 0\\
  1 & 0 & 0 & 0 & 0 & 0\\
  0 & 0 & 0 & 1 & 0 & 0\\
  0 & 1 & 0 & 0 & 0 & 0\\
  0 & 0 & 0 & 0 & 1 & 0\\
  0 & 0 & 1 & 0 & 0 & 0\\
  0 & 0 & 0 & 0 & 0 & 1\\

\end{pmatrix}\quad
\vec{\beta} =
\begin{pmatrix}
  x_1, \\ x_2, \\ x_3, \\ y_1, \\ y_2, \\ y_3
\end{pmatrix}
\vec{\gamma} =
\begin{pmatrix}
  dx_{1,2}, \\ dy_{1,2}, \\ dx_{2,3}, \\ dy_{2,3}, \\ -m_2gx_2 + gy_2,
  \\ gx_1, \\ gy_1, \\ tx_1, \\ ty_1, \\ tx_2, \\ ty_2, \\ tx_3, \\
  ty_3

\end{pmatrix}
\vec{W} =
\begin{pmatrix}
  1, \\ 1, \\ 1, \\ 1, \\ 1, \\ 1, \\ 1, \\ 0.01, \\ 0.01, \\ 0.01, \\
  0.01, \\ 0.01, \\ 0.01
\end{pmatrix}
\]

The variables to solve for are the $x_i$ and $y_i$ positions of
images, while equations are the feature-based distances between pairs
of images, images fixed to geometry with 0 or 1 degrees of freedom,
and the original noisy camera poses. In this scenario, a feature-based
distance of $dx_{1,2}$, $dy_{1,2}$ was calculated between images 1 and
2. This corresponds to the first and second row of $A$, while the
third and fourth row of $A$ represent the same for images 2 and
3. Rows 5 through 7 correspond to results of the geometry alignment
procedure in Section \ref{sec:geometryAlignment}. Specifically, row 5
corresponds to a geometry-based constraint of image 2's location to a
line of slope $m_2$, passing through point $gx_2$, $gy_2$, while rows
6 and 7 correspond to a fixed location for image 1 without any degrees
of freedom. Rows 8 through 13 correspond to the original camera
locations for each image ($tx_i,ty_i$).

The original camera poses are needed due to lack of feature matches in
all images, or lack of enough geometry alignment results to generate a
single solution. Since it is desirable to minimally use the original
noisy poses, we assign to them a weighting factor of $0.01$, while all
other equations are weighted at $1$.

This problem is linear, and it is solved quickly. After applying the
resulting shifts, images overlap and match each other with far greater
accuracy. Using the simple tile texturing scheme from Section
\ref{sec:simpleTextureMapping} on these adjusted images results in
Figure \ref{fig:compareAll}(b), which has far fewer discontinuities
than in \ref{fig:compareAll}(a), though some discontinuities remain
visible.

\section{Image Compositing}
\label{sec:imageCompositing}
In Section \ref{sec:simpleTextureMapping}, a simple tile-based image
selection and texturing method was described. After performing image
alignment, the image in Figure \ref{fig:compareAll}(b) was
obtained. It is clear that discontinuities in the image occur at tile
boundaries, and Section \ref{sec:mappingWithCaching} adds a caching
mechanism to reduce such boundaries. On the other hand, for special
cases where images have consistently perpendicular viewing
angles, Section \ref{sec:shortestPath} develops another image
compositing method that tends to produce cleaner textures. Both
methods are followed by a blending step, as shown in Figure
\ref{fig:flowchart}.

\subsection{Tile-Mapping with Caching}
\label{sec:mappingWithCaching}
To reduce discontinuities between tiles, this method reuses images for
adjacent tiles where possible, and otherwise selects images with
similar poses. This is accomplished via the use of a spatiotemporal
cache for selected images.

The best image for a tile $t$ is selected by searching through two
subsets of images for a viable candidate, before searching through the
entire set of valid images obtained in Section
\ref{sec:simpleTextureMapping}. The first subset of images is those
selected by adjacent tiles that have already been textured, and that
also contain texture for $t$. Of these images, one is selected
according to the scoring function in Figure
\ref{fig:scoringFunction}. Before reusing this image, we check the
criteria $\alpha < 45^\circ$, in order to ensure a high resolution
projection, with $\alpha$ as the camera angle as shown in Figure
\ref{fig:scoringFunction}.

If no satisfactory image is found in the first subset, a second subset
of images is used, consisting of those taken near the ones in the
first subset, both spatially and temporally. These images have quite
similar localization and projection, and are likely to match
well. If no viable image is found according to the $\alpha < 45^\circ$
criteria, we use the image selected from Section
\ref{sec:simpleTextureMapping}.

The result of applying this caching approach to the images for the
surface in Figure \ref{fig:compareAll}(a) is shown in Figure
\ref{fig:compareAll}(c), where seams are considerably reduced as
compared to Figure \ref{fig:compareAll}(b). However, some
discontinuities are still present, as visible in the posters on the
wall with breaks in their borders.

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=1in]{projectionWall.pdf}}
  \hspace{0.4in} \centering
  \subfloat[][]{\includegraphics[width=1.1in]{projectionCeiling.pdf}}
  \centering \hspace{0.4in}
  \subfloat[][]{\includegraphics[width=0.9in]{projectionWallCrop.pdf}}
  \caption{(a) Images for vertical planes are tilted, but their camera
    axes are more or less normal to their respective planes. (b)
    Camera axes for ceiling images are at large angles with respect to
    plane normals. (c) Wall images are cropped to be rectangular.}
  \label{fig:projectionAngles}
\end{figure}

\subsection{Shortest Path Texturing}
\label{sec:shortestPath}

As mentioned earlier, our data comes from a mobile backpack system
carried by am ambulatory human operator, typically bent forwards at 15
to 20 degrees with respect to the vertical direction. As a result,
cameras facing sideways are head on with respect to vertical walls, as
shown in Figure \ref{fig:projectionAngles}(a), while cameras oriented
towards the top or bottom of the backpack are at an angle with respect
to horizontal floors and ceilings, as shown in Figure
\ref{fig:projectionAngles}(b). These oblique camera angles for
horizontal surfaces translate into textures that span large areas once
projected, as shown in Figure \ref{fig:projectionAngles}(b). Using the
tile-based texture mapping criteria from Figure
\ref{fig:scoringFunction}, such projections have highly varying scores
depending on the location of a tile on the plane. Thus, the tiling
approach in Section \ref{sec:mappingWithCaching} is an appropriate
choice for texturing floors and ceilings, as it uses the parts of
image projections that maximize resolution for their respective plane
locations, e.g. areas near point A and not near point B, in Figure
\ref{fig:projectionAngles}(b).

For vertical walls however, most images are taken from close distances
and head-on angles, resulting in high resolution fronto-parallel
projections. As a result, for each tile on a wall plane, the scoring
function of Figure \ref{fig:scoringFunction} is relatively flat with
respect to candidate images, as they are all more or less head
on. Since the scoring function is less discriminative for walls, it is
conceivable to devise a different texturing strategy to directly
minimize visible seams when texturing them. This is done by choosing
the smallest possible subset of images from the set selected in
Section \ref{sec:simpleTextureMapping} and aligned in Section
\ref{sec:2dAlignment} such that it (a) covers the entire plane and (b)
minimizes the visibility of borders between the images. A
straightforward cost function that accomplishes the latter is the sum
of squared differences (SSD) of pixels in overlapping regions between
all pairs of images. Minimizing this cost function encourages image
boundaries to occur either in featureless areas, such as bare walls,
or in areas where images match extremely well.

In its most general form, our problem can be defined as minimally
covering a polygon i.e. the planar surface, using other polygons of
arbitrary geometry i.e. image projections, with the added constraint
of minimizing the cost function between chosen images. Given that
wall-texture candidate images are taken from more or less head-on
angles, and knowing that generally minor rotations are made in Section
\ref{sec:2dAlignment}, we can crop our image projections to be
rectangular with minimal texture loss, as shown in Figure
\ref{fig:projectionAngles}(c). Furthermore, because the fisheye camera
lenses have full floor-to-ceiling coverage of nearly all walls, and
the backpack operator nearly always only moves horizontally, we only
need to ensure lateral coverage of our wall planes. We can thus
construct a Directed Acyclic Graph (DAG) from the images, with edge
costs defined by the SSD cost function, and solve a simple shortest
path problem to find an optimal subset of images with regard to the
SSD cost function \cite{dijkstra}.

\begin{figure}
  \centering
  \includegraphics[width=4.8in]{dagCreation.pdf}
  \caption{DAG construction for the image selection process. \\}
  \label{fig:dagCreation}
\end{figure}

Figure \ref{fig:dagCreation} demonstrates the construction of a DAG
from overlapping images of a hallway wall. Images are nodes in a
graph, and directed edges are placed in the graph from left to right
between overlapping images. The weights of these edges are determined
by the SSD cost function. Next, two artificial nodes representing the
ends of the plane are added, with equal-cost $C_0$ edges to the images
crossing them. We now solve the shortest path problem from one end to
the other. This results in a set of images completely covering the
plane horizontally, while minimizing the cost of seams between images.

As seen in Figure \ref{fig:compareAll}(d), this shortest path method
has fewer visible discontinuities than Figure \ref{fig:compareAll}(c)
corresponding to the tile caching approach\footnote{In Figure
  \ref{fig:compareAll}(d), we arbitrarily chose one image for
  texturing where images overlap, as blending will be discussed in
  section \ref{sec:blending}.}. This is especially evident when
comparing the posters in the images. This shortest path approach
directly reduces the cost of each image boundary, while the tile
caching method uses a scoring function that only approximates this
effect. Furthermore, this approach guarantees the best selection of
images to minimize seams, while the sequential tile caching method may
select images early on that turn out to be poor choices once
subsequent tiles have been processed.

When texturing an entire 3D planar model, we apply the shortest path
method on walls, due to its superior results when provided with
head-on images with lateral coverage. Floors and ceilings however,
given their many images taken at oblique angles, are textured using
the tile caching method of Section \ref{sec:mappingWithCaching}.

\subsection{Blending}
\label{sec:blending}
We now apply a blending procedure to the texturing methods in Sections
\ref{sec:mappingWithCaching} and \ref{sec:shortestPath}. Although the
image alignment steps and image selection in both methods attempt to
minimize all mismatches between images, there are occasional
unavoidable discontinuities in the final texture due to different
lighting conditions or inaccuracies in model geometry. These can
however be treated and smoothed over by blending over image seams.
Whether the units to be blended are rectangularly-cropped images or
rectangular tiles, we can apply the same blending procedure, as long
as there is a guaranteed overlap between images.

For the tile caching method of Section \ref{sec:mappingWithCaching},
overlap can be ensured by texturing a larger tile than needed for
display. For example, for a rendered tile $l_1 \times l_1$, we can
associate it with a texture $(l_1 + l_2) \times (l_1 + l_2)$ in size.
For the shortest path method, overlap between images is inherently
ensured. To enforce consistent blending however, a minimum required
overlap between images of 200 mm for the graph is added.

After linear alpha blending across overlapping regions, the texture
mapping process is complete. Figures \ref{fig:compareAll}(e) and
\ref{fig:compareAll}(f) show the blended versions of Figures
\ref{fig:compareAll}(c) and \ref{fig:compareAll}(d) respectively. The
remaining images in Figure \ref{fig:compareAll} highlight differences
between the two methods, showing that Figure \ref{fig:compareAll}(f)
has the cleanest texture and contains the best visual quality among
the textures in Figure \ref{fig:compareAll}.

\section{Results}
\label{sec:results}
Examples of ceilings and floors textured with the tile caching
approach, and walls textured with the shortest path approach, are
displayed in Figure \ref{fig:results}. High resolution colored texture
comparisons, as well as video and interactive walkthroughs of full
models are available at
\footnote{\url{http://www-video.eecs.berkeley.edu/research/indoor/}}.

As mentioned earlier, our approach is quite efficient. The top wall in
Figure \ref{fig:results}(a) was generated with 7543 $\times$ 776
pixels, and spans a 40-meter long wall. Given 41000 input images of
the entire dataset, a 2.8GHz dual-core consumer-grade laptop takes
approximately a minute to choose 36 candidate images, followed by
under a minute to perform both image alignment and the shortest path
texturing method, though over 75\% of that time is spent calculating
SIFT matches within the SiftGPU framework, which could feasibly be
split into a separate preprocessing step. While not real-time, the
process is capable of generating fast updates after changes in various
parameters or modifications to input data, and if integrated directly
into a 3D modeling system, could provide live visualization as data is
collected. Our full models consist of an input model file, textures,
and a mapping of image points to 3D model vertices. The models shown
in Figure \ref{fig:results} are roughly 20 MB in size, and are
visualized using the OpenSceneGraph toolkit \cite{openscenegraph},
which allows for export to many common model formats, as well as
efficient visualization, even in web browsers or mobile devices.

\section{Conclusion}
\label{sec:conclusion}

In this paper, we have developed an approach to texture mapping models
with noisy camera localization data. We are able to refine image
locations based on geometry references and feature matching, and
robustly handle outliers. Using the tile-based mapping approach, we
can texture both simple rectangular walls as well as complex floor and
ceiling geometry. We also implemented a shortest path texturing method
that produces seamless textures on planes where multiple head-on
images are available. Both of these approaches are highly modular, and
easily tunable for similar systems across multiple environments.

Our method is likely to fail in scenarios where 3D error is large. A
logical progression of our approach to resolve camera error in 3D is
to perform matching between image lines and geometry in 3D, which can
be done reasonably efficiently \cite{linebased,
  rectangularstructures}. Using linear features in addition to SIFT
features is also likely to result in improved matches, as indoor
scenes often have long, unbroken lines spanning multiple images
\cite{linearposeestimation}. Finally, the blending procedure is quite
basic, and applying more sophisticated methods of blending as well as
normalization would benefit the final visual quality, and more
robustly handle motion-based or parallax errors.

\begin{figure}
  \centering
  \subfloat[][]{\includegraphics[width=2.4in]{finalfloors.jpg}} ~~~~~~~~
  \centering
  \subfloat[][]{\includegraphics[width=2.4in]{finalceilings.jpg}}

  \centering \subfloat[][]{
    \includegraphics[height=2in, width=2.3in]{floorcropped.jpg} ~~~~~~~~
    \includegraphics[width=2.3in]{pier15floor.jpg}
  }

  \centering{
    \includegraphics[width=2.3in]{fullmodel.png}
    \includegraphics[width=2.3in]{pier15.png} \\
    \includegraphics[width=2.3in]{threestoryfull2.png}
    \includegraphics[width=2.3in]{threestoryfull.png}
  }

  \centering \subfloat[][]{ }
  \caption{Examples of our final texture mapping output for (a) walls,
    (b) ceilings, (c) floors, (d) full models.}
  \label{fig:results}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%% References %%%%%

\bibliography{report} %>>>> bibliography data in report.bib
\bibliographystyle{spiebib} %>>>> makes bibtex use spiebib.bst

\end{document}